### 2018-09-05 src/biological_data_prep.Rmd
Solve missinginstitution issue

---
title: "biological data preperation"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Preperation of the capelin (*Mallotus villosus*) data obtained from [OBIS](www.iobis.org). This database shows the global occurance of capelin. 
Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("packrat", "rworldmap")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

# Raw Data Clean
Load the raw data, then check use head() to check it looks ok, then names() to see all the column names
```{r load raw data}
rawdata <- read.csv("../data/bio/capelin_obis_raw.csv")
head(rawdata) 
names(rawdata)
```

following [Darwin Core](http://rs.tdwg.org/dwc/terms/) rename the latitude and longitude columns to decimalLatitude and decimalLongitude respectively
```{r dawin core naming convention}
tidydata <- rawdata
names(tidydata)[names(tidydata)=="latitude"] <- "decimalLatitude"
names(tidydata)[names(tidydata)=="longitude"] <- "decimalLongitude"
head(tidydata)
```

The dataset is global and needs to be trimmed to contain just the area of interest. 
Plot all the occurance points
```{r map all occurrence points}
library(rworldmap)
map <- getMap(resolution = "low") #creates an object called map at medium resoultion
plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) #the x and y lim are the long-lat bounds of the map
points(tidydata$decimalLongitude, tidydata$decimalLatitude) #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```
now extract just the Atlantic Canada data.
The lon-lat study area (which captures the whole Atlantic Canadian EEZ up to the north of Resolution Island in the Hudson Strait/ Davis Strait) is -71.0,-43.0,62.0,38.0 
```{r map only atlantic canadian occurrences}
tidydata <- tidydata[tidydata$decimalLongitude > -71, ]
tidydata <- tidydata[tidydata$decimalLongitude < -43, ]
tidydata <- tidydata[tidydata$decimalLatitude > 38, ]
plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) #re-map - this is just atl. canada (after checked subsetting is good)
points(tidydata$decimalLongitude, tidydata$decimalLatitude)
```
ok looks good...save what you have so far to a .csv
```{r save tidydata to csv}
write.csv(tidydata, file = "../data//bio/capelin_obis_tidy.csv", row.names = FALSE)
```

# Column removal
So there are many columns that you don't need.. you want to keep
id
decimalLatitude
decimalLongitude
datecollected
institutioncode
individualcount
resname
originalscientificname

```{r select colums needed}
tidydata <- subset(tidydata, select = c(id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, resname, originalscientificname))
head(tidydata)
```

# remove data without month/year info

count NAs in the dataset. Mainly you are interested in the datecollected column (ignore individual count - you know there is a lot missing)
```{r}
missingdata <- apply(tidydata, 2, function (x) sum(is.na(x)))
missingdata
```
hmm interesting - you have 771 columns from year month and day with missing data, but datacollected has 0. There may be a formatting issue.
To take a closer look, subset all rows where year == na

```{r}
missingdatarows <- tidydata[is.na(tidydata$year), ]
head(missingdatarows)
```
so the NAs in the year column relate to no data (without a na value attached). Remove all rows where year == na
```{r}
tidydata <- tidydata[!is.na(tidydata$year), ]
missingdata2 <- apply(tidydata, 2, function (x) sum(is.na(x)))
missingdata2
```

# Month year column creation
make columns for the years and month based on the datecollected column
```{r new columns}
tidydata$year <- format(as.Date(tidydata$datecollected), "%Y") #created a new column called year
tidydata$month <- format(as.Date(tidydata$datecollected), "%m") #created a new column called month
tidydata$day <- format(as.Date(tidydata$datecollected), "%d") #created a new column called day 
head(tidydata)
```

the year, month, day columns are characters Change to numeric
```{r new columns 2}
tidydata$year <- as.numeric(tidydata$year)
tidydata$month <- as.numeric(tidydata$month)
tidydata$day <- as.numeric(tidydata$day)
head(tidydata)
```

# Institutions
Lets take a look at the institutioncode to make sure the names make sense...
```{r}
tidydata$institutioncode <- as.character(tidydata$institutioncode)
whosampled <- unique(tidydata$institutioncode) #makes a variable of institutions
whosampledordered <- sort(whosampled) #just puts the list in 
whosampledordered
```
Ok so a number of these names actually represent the same insitiution (e.g. [DFO Regions](http://www.dfo-mpo.gc.ca/regions/index-eng.htm). Some details [here](ftp://ftp.aoml.noaa.gov/od/pub/library/methods-branton.pdf) and the first is no institution (you might be able to figure it out though). 
The merges needed are as follows
"Bedford Institute of Oceanography (BIO)" & "BIO" & "DFO-SF" == DFOMTMS 
"DFO-IML" & "IML-MLI" & "IML_MLI" & "DFO-NG" == DFOQC 
"DFO-NFLD" & "North Atlantic Fisheries Centre" == "DFONL" 
"NMFS-NEFSC" & "NOAA, NMFS, Northeast Fisheries Science Center" == NEFSC
"The Atlantic reference Centre" & "The Atlantic Reference Centre (ARC)" == ARC 
"DFO-SG" & "DFO Gulf Fisheries Centre" == DFOGULF

and some you can just make shorter
"Northwest Atlantic Fisheries Organization (NAFO)" == NAFO
"DFO-ISDM" == DFOISDM
"Maine DMR" == MaineDMR
```{r}
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "Bedford Institute of Oceanography (BIO)", "DFOMTMS")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "BIO", "DFOMTMS")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-SF", "DFOMTMS")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-IML", "DFOQC")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "IML-MLI", "DFOQC")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-NG", "DFOQC")  
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "IML_MLI", "DFOQC")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-NFLD", "DFONL")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "North Atlantic Fisheries Centre", "DFONL")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "NMFS-NEFSC", "NEFSC")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "NOAA, NMFS, Northeast Fisheries Science Center", "NEFSC")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "The Atlantic Reference Centre (ARC)", "ARC")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "The Atlantic reference Centre", "ARC")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-SG", "DFOGulf")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO Gulf Fisheries Centre", "DFOGulf")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "Northwest Atlantic Fisheries Organization (NAFO)", "NAFO")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-Central and Arctic, Freshwater Institute (FWI)", "DFOCENARC")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-ISDM", "DFOISDM")
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "Maine DMR", "MaineDMR")

#recheck
tidydata$institutioncode <- as.character(tidydata$institutioncode)
whosampled <- unique(tidydata$institutioncode) #makes a variable of institutions
whosampledordered <- sort(whosampled) #just puts the list in 
whosampledordered
```

Now look at the missing institution. Use resname to see if it helps...
```{r}
missinginst <- subset(tidydata, institutioncode == "")
missinginst
```
ok now check for unique resname...
```{r}
missinginst$resname <- as.character(missinginst$resname)
missinginstunique <- unique(missinginst$resname) #makes a variable of institutions
missinginstuniquedordered <- sort(missinginstunique) #just puts the list in 
missinginstuniquedordered <- as.character(missinginstuniquedordered)
missinginstuniquedordered
```

Ok soo looking at the [OBIS metadata](http://www.iobis.org/explore/#/dataset/4356), the institution is the Conservation of Arctic Flora and Fauna (CAFF)
```{r}
tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "", "CAFF")
whosampled <- unique(tidydata$institutioncode) #makes a variable of institutions
whosampledordered <- sort(whosampled) #just puts the list in 
whosampledordered
```