### 2018-09-28 src/raster_brick_stuff.Rmd
probably not useful stuff for my netcdf data extraction...

---
title: "raster_brick_stuff"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Creating a way to automate mass raster_brick creation, with unique names...

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. The code herein is unlikely to be elegant and there is likely more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first. If they are not installed, it will do so before loading
```{r pre-install packages, suppressMessages(expr)}
packages <- c("rgdal", "raster", "ncdf4", "RNetCDF", "sp", "parallel", "magrittr", "data.table") 
source("../src/ipak.R")
ipak(packages)
```

For the parallel computing (this will make things faster), detect cores automatically (leave one free though).
```{r}
cores <- detectCores() - 1
```

Anne is the best - really the best. She is the best that ever was. No Anne bester.

Tell R where to find your data and where to spit out the output
```{r data_paths}
envdata_path <- "../data/env/netcdf"
projected_path <- "../data/env"
output_path <- "../output/env"
```