### 2018-10-02 src/raster_brick_stuff.Rmd
more raster loop mess...

---
title: "raster_brick_stuff"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Creating a way to automate mass raster_brick creation, with unique names...

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. The code herein is unlikely to be elegant and there is likely more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first. If they are not installed, it will do so before loading
```{r pre-install packages, suppressMessages(expr)}
packages <- c("rgdal", "raster", "ncdf4", "RNetCDF", "sp", "parallel", "magrittr", "data.table") 
source("../src/ipak.R")
ipak(packages)
```

Note that the original files have been renamed using Windows Powershell. See OneNote document for coding details. All files are now in the following format:

yyyy_mm_variablename.nc

where variablename is all alpha characters, but varies in length. E.G. 1993_01_mlp.nc

For the parallel computing (this will make things faster), detect cores automatically (leave one free though).
```{r}
cores <- detectCores() - 1
```

Tell R where to find your data and where to spit out the output
```{r data_paths}
weatherraster_path <- "../data/env/netcdftest"
countryshape_path <- "./output/env"
output_path <- "../output/env/nctest"
```

# ---- 1. Define Functions.

# -- 1.A. Define Small Subfunctions.

Small function 1) Reads filename & explicitly opens it as a NetCDF file.
```{r}
open_netcdf_as_rasterbrick <- function( ncdf_filename_input ) {
  ncdf_filename_input %>%
  file.path( weatherraster_path , . ) %>%
  nc_open( . )  %>%  # Open path as NetCDF file.
  ncvar_get( . ) %>%  # Get NetCDF file.
  brick( . ) %>%
  return( . )
}
```

# Small function 2) Transforms the raster brick to our country shapefile.
```{r}
match_rainbrick_to_countryshape <- function( brick_input ) {
  brick_input %>%
  # Reproject raster brick to the shapefile's coordinate system.
  projectRaster( . , crs = proj4string( country_shapefile ), method = "ngb" ) %>%
  return( . )
}
```

# Small function 3) Extract data from a raster brick.
```{r}
generate_data_from_rasterandshape <- function( brick_input ) {
	brick_input %>%
    raster::extract( . , countryshape , TRUE, fun = mean, na.rm = TRUE  )
    return( . )
}
```

# Small function 4) Grab 4-digit year from input filename.
```{r}
grab_year_from_inputfile <- function( ncdf_filename_input ) {
  ncdf_filename_input %>%
  regexpr("[0-9]+", . ) %>%  # Match 4-digit year.
  regmatches( ncdf_filename_input , . ) %>%  # Get matched REGEX from input string.
  return( . )
}
```

# --- 1.B. Define "BIG" Function That Extracts Dataset From a NetCDF File.
```{r}
generate_datatable_from_rasterbricks <- function( ncdf_filename_input ) {
	ncdf_filename_input %>%
	open_netcdf_as_rasterbrick( . ) %>%
	match_rainbrick_to_countryshape( . ) %>%
	generate_data_from_rasterandshape( . ) -> country_means_dataframe
	ncdf_filename_input %>%
	grab_year_from_inputfile( ) %>%
	write.csv( country_means_dataframe ,
				file = file.path( . , output_path ) )
}
```

# ---- 2. Main Code: Setup Environment to Run Big Function.

Start with your name of the country shapefile we're referencing.
```{r}
"poly.shp" %>%
file.path(countryshape_path) %>%
readODG( den = .  , layer = "FID" ) -> countryshape
raster_file_list <- list.files( path = weatherraster_path ,
pattern = ".nc" , all.files = FALSE , full.names = FALSE )
```

Generate list of NetCDF files automatically from our directory.
Match all files ending in ".nc"
```{r}
raster_file_list <- list.files( path = weatherraster_path ,
pattern = ".nc" , all.files = FALSE , full.names = FALSE )
```

Run our big function on the list of NetCDF files.
```{r}
mclapply( raster_file_list , generate_datatable_from_rasterbricks )
```

# ---- 3. Assemble .CSV Files using Data.Table and Lapply.

Fetch all files ending in .CSV in out output path.
```{r}
csv_file_list <- list.files( path = output_path ,
                             pattern = ".csv",
                             all.files = FALSE,
                             full.names = TRUE,
                             recursive = FALSE )
lapply( csv_file_list , fread , sep = "," ) %>%
rbindlist( . ) -> big_datatable
write.csv( big_datatable, file = file.path( output_path , "big_file_name.csv") )
```

# !!!!!

```{r}
b <- nc_open("../data/env/netcdftest/1993_01_salinity.nc")
b
c <- brick(b, lvar = 4)
c
pts <- data_aea
ts <- extract(c, cbind(pts$decimalLongitude, pts$decimalLatitude))
```