### 2018-10-03 src/help_extract_from_multiple_netCDFs.Rmd
seeking extraction help - tidy code to do this

---
title: "help_extract_from_multiple_netCDFs"
author: "Samantha Andrews"
output: html_notebook
---

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first. If they are not installed, it will do so before loading
```{r pre-install packages, suppressMessages(expr)}
packages <- c("rgdal", "raster", "ncdf4", "RNetCDF", "sp", "parallel", "magrittr", "data.table") 
source("../src/ipak.R")
ipak(packages)
```

# Initial setup

Detect cores automatically, I usually free one up.
```{r}
cores <- detectCores() - 1
```

Define your file paths here.
```{r}
weatherraster_path <- "/path/to/weatherrasterfiles"
countryshape_path <- "/path/to/countryshapefile"
output_path <- "/path/to/outputfiles"
```

# ---- 1. Define Functions.

# -- 1.A. Define Small Subfunctions.

Small function 1) Reads filename & explicitly opens it as a NetCDF file.
```{r}
open_netcdf_as_rasterbrick <- function( ncdf_filename_input ) {
  ncdf_filename_input %>%
  file.path( weatherraster_path , . ) %>%
  nc_open( . )  %>%  
  ncvar_get( . ) %>%   
  brick( . ) %>% # original code: brick( . ,ymn = -0, ymx = 360, xmn = -90, xmx = 90, crs = "the string for your CRS" ) %>%
  return( . )
}
```

Small function 2) Transforms the raster brick to our country shapefile.
```{r}
match_rainbrick_to_countryshape <- function( brick_input ) {
  brick_input %>%
  projectRaster( . , aea) %>% #aea represents an existing rasterbrick that is in the albers equal area   projection
  # Original code included a crop option here - my files are already cropped to the region I need
  return( . )
}
```

# Small function 3) Extract data from a raster brick.
```{r}
generate_data_from_rasterandshape <- function( brick_input ) {
	brick_input %>%
	# Take means according to the countryshape.
	# Make sure df = TRUE , so that output is a dataframe.
    raster::extract( . , countryshape ,
    				f=TRUE, fun = mean, na.rm = TRUE  )
    return( . )
}
```

# Small function 4) Grab 4-digit year from input filename.
grab_year_from_inputfile <- function( ncdf_filename_input ) {

  ncdf_filename_input %>%
  regexpr("[0-9]+", . ) %>%  # Match 4-digit year.
  regmatches( ncdf_filename_input , . ) %>%  # Get matched REGEX from input string.
  return( . )
}

# --- 1.B. Define "BIG" Function That Extracts Dataset From a NetCDF File.

generate_datatable_from_rasterbricks <- function( ncdf_filename_input ) {

	# Note: the only argument is a NetCDF filename.

	# Start with file argument and process with the sub-functions above.
	ncdf_filename_input %>%

	open_netcdf_as_rasterbrick( . ) %>%
	match_rainbrick_to_countryshape( . ) %>%
	generate_data_from_rasterandshape( . ) -> country_means_dataframe

	# Go back to the file input name, create automatic names, and save.
	ncdf_filename_input %>%

	grab_year_from_inputfile( ) %>%
	write.csv( country_means_dataframe ,
				file = file.path( . , output_path ) )
}

# ---- 2. Main Code: Setup Environment to Run Big Function.

# Start with your name of the country shapefile we're referencing.
"country_shapefile_name.shp" %>%
file.path( countryshape_path , ) %>%
readODG( den = .  , layer = "countries" ) -> countryshape

# Generate list of NetCDF files automatically from our directory.
# Match all files ending in ".nc"
raster_file_list <- list.files( path = weatherraster_path ,
pattern = ".nc" , all.files = FALSE , full.names = FALSE )

# Run our big function on the list of NetCDF files.
mclapply( raster_file_list , generate_datatable_from_rasterbricks )

# ---- 3. Assemble .CSV Files using Data.Table and Lapply.

# Fetch all files ending in .CSV in out output path.
csv_file_list <- list.files( path = output_path ,
                             pattern = ".csv",
                             all.files = FALSE,
                             full.names = TRUE,
                             recursive = FALSE )

# Take the list of saved files & "fast read" them into R.
lapply( csv_file_list , fread , sep = "," ) %>%

# Transform the list of read files into a data.table:
rbindlist( . ) -> big_datatable

# Note: Before reassembling the data, or after, you may want
# to manipulate the data so that it is in a more usable format.

# Note: You may want to setkeys() for data.table here.

# Save the big file.
write.csv( big_datatable, file = file.path( output_path , "big_file_name.csv") )