### 2019-03-05 src/stack_test.Rmd
ascii-file database

This works

need raster rgdal netcdf4 tidyr

```{r}
ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
rasters <- lapply(paste0(ascname), raster)

year <- "X1998"
month <- "01"
depth <- "86.9294"
tempvar <- "temp"
chlvar <- "chl"
o2var <- "o2"
salvar <- "salinity"
depthvar <- "depth"
surfacevar <- "surface"
naovar <- "nao"
amovar <- "amo"
samplevar <- "sample"
wintervar <- "winter"

for (i in 1:length(rasters)){
  ascyear <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 1))
  ascmonth <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 2))
  ascdepth <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 5))
  ascvar1 <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 3))
  ascvar2 <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 4))
  ascfile <- rasters[i]

    if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
    #print(ascfile)
    temp_depth <- ascfile[[1]]
    names(temp_depth) <- "temp_depth"
  } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
    chl_surface <- ascfile[[1]]
    names(chl_surface) <- "chl_surface"
  } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
    o2_depth <- ascfile[[1]]
    names(o2_depth) <- "o2_depth"
  } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
    salinity_depth <- ascfile[[1]]
    names(salinity_depth) <- "salinity_depth"
  } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == wintervar){
    nao_winter <- ascfile[[1]]
    names(nao_winter) <- "nao_winter"
  } else if (ascyear == year & ascmonth == month  & ascvar1 == naovar & ascvar2 == samplevar){
    nao_sample <- ascfile[[1]]
    names(nao_sample) <- "nao_sample"

  }
   st19980186.9 <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
  }
st19980186.9
```

messing with code to see if can do for unique year month depth

need tidyr package
```{r}
unyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
unmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
undepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85", "1151.99", "1265.86", "1387.38") #this list is based on how the files are named when split into depth layers. note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers

unlayer <- crossing(unyears, unmonths, undepths)
unlayer$stackname <- paste0("st", unlayer$unyears, unlayer$unmonths, unlayer$undepths) #this creates the name of the stack that will be used
unlayer$temp_depth <- NA
unlayer$temp_surface <- NA
unlayer$chl_depth <- NA
unlayer$chl_surface <- NA
unlayer$o2_depth <- NA
unlayer$o2_surface <- NA
unlayer$salinity_depth <- NA
unlayer$salinity_surface <- NA
unlayer$nao_sample <- NA
unlayer$nao_prev <- NA
unlayer$nao_winter <- NA
unlayer$amo_sample <- NA
unlayer$amo_prev <- NA
unlayer$amo_winter <- NA

ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

#year <- "X1998"
#month <- "01"
#depth <- "86.9294"
tempvar <- "temp"
chlvar <- "chl"
o2var <- "o2"
salvar <- "salinity"
depthvar <- "depth"
surfacevar <- "surface"
naovar <- "nao"
amovar <- "amo"
samplevar <- "sample"
wintervar <- "winter"

for (k in 1:nrow(unlayer)){
    print(unlayer$stackname[k])
    year <- unlayer$unyears[k]
    month <- unlayer$unmonths[k]
    depth <- unlayer$undepths[k]
    for (i in 1:length(ascname)){

    ascyear <- (sapply(strsplit(ascname[[i]], "_"), "[[", 1))
    ascmonth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 2))
    ascdepth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 5))
    ascdepth <- gsub(".asc", "", ascdepth)
    ascvar1 <- (sapply(strsplit(ascname[[i]], "_"), "[[", 3))
    ascvar2 <- (sapply(strsplit(ascname[[i]], "_"), "[[", 4))
    ascfile <- ascpath[[i]]
    print(ascfile)

    if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
    unlayer$temp_depth[k] <- ascfile
  } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
   unlayer$chl_surface[k] <- ascfile
 } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
  unlayer$o2_depth[k] <- ascfile
 } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
  unlayer$salinity_depth[k] <- ascfile
 } else if (ascyear == year & ascvar1 == naovar & ascvar2 == wintervar){
  unlayer$nao_winter[k] <- ascfile
 } else if (ascyear == year & ascmonth == month  & ascvar1 == naovar & ascvar2 == samplevar){
  unlayer$nao_sample[k] <- ascfile
 }

    }
    write.csv(unlayer, "../data/env/asc_layerlist.csv", row.names = TRUE)
}

# now loop to create stacks
unlayer <- read.csv("../data/env/asc_layerlist.csv", header = TRUE)
unlayer[] <- lapply(unlayer, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
aea <- raster("../output/env/aea.tif")

for (td in 1:nrow(unlayer)){
  temp_depth <- raster(unlayer$temp_depth[td])
  #crs(temp_depth) <- +proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 
  names(temp_depth) <- "temp_depth"
  #temp_depth <- projectRaster(temp_depth, aea)
  chl_surface <- raster(unlayer$chl_surface[td])
  names(chl_surface) <- "chl_surface"
  o2_depth <- raster(unlayer$o2_depth[td])
  names(o2_depth) <- "o2_depth"
  salinity_depth <- raster(unlayer$salinity_depth[td])
  names(salinity_depth) <- "salinity_depth"
  nao_sample <- raster(unlayer$nao_sample[td])
  names(nao_sample) <- "nao_sample"
  nao_winter <- raster(unlayer$nao_winter[td])
  names(nao_winter) <- "nao_winter"
  stackn <- unlayer$stackname[td]
  test <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
  #stackSave(test,unlayer$stackname[td])
  pred <- predict(mod1, test, file = paste0("../output/maxentpredict/", stackn), overwrite=TRUE) 
  preplot <- plot(pred, png(paste0("../output/maxentpredict/", stackn, ".png")))
  paste("maxplot_", stackn, ".png", sep = "") # to automatically save the plot to a png AND show it inline

}
dev.off() # stops automatic saving of the plot to a png
# now call the stack...
{
stack <- stackOpen("st19920310.stk")
}
```

  preplot <- plot(pred, png(paste0(stackn, ".png")))

```{r}
var <- raster(stacktest$chl[1])
var
a <- raster(var)
a
```

#maxent
maxnet throwing errors...
```{r}
janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
janpres$occ <- 1
janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
janab$occ <- 0
prab <- rbind(janpres, janab)
prab$originalscientificname <- "cap" #because species name needs to be the same
pa <- prab$occ
prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
prab

system.file("java", package="dismo") #finding location to put maxent jar file
system("java -version")
mod1 <- maxent(x = prab, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))
pre1 <- predict(mod1, test) #,  args='outputformat=cloglog'
plot(pre1)
plot(mod1)

o2_depth <- raster("../data/jan_test_86.9/2005_01_o2.nc_0.50576.asc")
names(o2_depth) <- "o2_depth"

test2 <- stack(temp_depth, salinity_depth, o2_depth, chl_surface)
prab2 <- subset(prab, select = -c(nao_sample, nao_winter)) #ok so for maxent in r just need the variables
mod2 <- maxent(x = prab2, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))
pre2 <- predict(mod2, test2)
plot(pre2)
plot(pre1)
```

```{r}
unyears <- as.character(c("1992", "1998")) 
unmonths <- c("01", "03")
undepths <- c("10", "86.9294")
unlayer <- crossing(unyears, unmonths, undepths)
unlayer$stackname <- paste0("st", unlayer$unyears, unlayer$unmonths, unlayer$undepths) #this creates the name of the stack that will be used

unlayer$temp_depth <- NA
unlayer$temp_surface <- NA
unlayer$chl_depth <- NA
unlayer$chl_surface <- NA
unlayer$o2_depth <- NA
unlayer$o2_surface <- NA
unlayer$salinity_depth <- NA
unlayer$salinity_surface <- NA
unlayer$nao_sample <- NA
unlayer$nao_prev <- NA
unlayer$nao_winter <- NA
unlayer$amo_sample <- NA
unlayer$amo_prev <- NA
unlayer$amo_winter <- NA

tempvar <- "temp"
chlvar <- "chl"
o2var <- "o2"
salvar <- "salinity"
depthvar <- "depth"
surfacevar <- "0.50576"
naovar <- "nao"
amovar <- "amo"
samplevar <- "sample"
wintervar <- "winter"

ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

for (k in 1:nrow(unlayer)){
    year <- unlayer$unyears[k]
    month <- unlayer$unmonths[k]
    depth <- unlayer$undepths[k]

    for (i in 1:length(ascname)){
    ascyear <- (sapply(strsplit(ascname[[i]], "_"), "[[", 1))
    ascmonth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 2))
    ascvar <- (sapply(strsplit(ascname[[i]], "_"), "[[", 3))
    ascdepth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 4))
    ascdepth <- gsub(".asc", "", ascdepth)
    ascfile <- ascpath[[i]]
    print(ascfile)
    if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
      unlayer$temp_depth[k] <- ascfile
    } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
        unlayer$temp_surface[i] <- ascfile
    }
    }
}

```