### 2019-03-07 src/asc_file_database.Rmd
arghhhhh!!!

---
title: "asc_file_database"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Preperation database to extract and stack asc files

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there are probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.

```{r pre-install packages, message=FALSE}
packages <- c("tidyr")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

```{r}
ascyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
ascmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
ascdepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85", "1151.99", "1265.86", "1387.38") 
#this depth list is based on how the files are named when split into depth layers. note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers

asclayerall <- crossing(ascyears, ascmonths, ascdepths)
asclayerall$stackname <- paste0("st", asclayerall$ascyears, asclayerall$ascmonths, asclayerall$ascdepths) #this creates the name of the stack that will be used
asclayerall$temp_depth <- NA
asclayerall$temp_surface <- NA
asclayerall$chl_depth <- NA
asclayerall$chl_surface <- NA
asclayerall$o2_depth <- NA
asclayerall$o2_surface <- NA
asclayerall$salinity_depth <- NA
asclayerall$salinity_surface <- NA
asclayerall$ssh_surface <- NA
asclayerall$mlp_surface <- NA
asclayerall$nao_sample <- NA
asclayerall$nao_prev <- NA
asclayerall$nao_winter <- NA
asclayerall$amo_sample <- NA
asclayerall$amo_prev <- NA
asclayerall$amo_winter <- NA

```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 1998)

ascname <- list.files("../data/env/ascii/ascii1998/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii1998/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist1998.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCLAYER ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 1999)

ascname <- list.files("../data/env/ascii/ascii1999/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii1999/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist1999.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2000)

ascname <- list.files("../data/env/ascii/ascii2000/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2000/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2000.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2001)

ascname <- list.files("../data/env/ascii/ascii2001/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2001/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2001.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2002)

ascname <- list.files("../data/env/ascii/ascii2002/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2002/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2002.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2003)

ascname <- list.files("../data/env/ascii/ascii2003/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2003/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2003.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2004)

ascname <- list.files("../data/env/ascii/ascii2004/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2004/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2004.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2005)

ascname <- list.files("../data/env/ascii/ascii2005/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2005/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2005.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2006)

ascname <- list.files("../data/env/ascii/ascii2006/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2006/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2006.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2007)

ascname <- list.files("../data/env/ascii/ascii2007/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2007/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2007.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2008)

ascname <- list.files("../data/env/ascii/ascii2008/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2008/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2008.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2009)

ascname <- list.files("../data/env/ascii/ascii2009/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2009/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2009.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2010)

ascname <- list.files("../data/env/ascii/ascii2010/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2010/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2010.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2011)

ascname <- list.files("../data/env/ascii/ascii2011/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2011/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2011.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2012)

ascname <- list.files("../data/env/ascii/ascii2012/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2012/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2012.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2103)

ascname <- list.files("../data/env/ascii/ascii2013/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2013/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2013.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2014)

ascname <- list.files("../data/env/ascii/ascii2014/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2014/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2014.csv", row.names = TRUE)
}
```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2015)

ascname <- list.files("../data/env/ascii/ascii2015/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/ascii2015/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.nc"
chlvar <- "chl.nc"
o2var <- "o2.nc"
salvar <- "salinity.nc"
mlpvar <- "mlp.nc"
sshvar <- "ssh.nc"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprevvar <- "naoprev"
amoprevvar <- "amoprev"
naowintervar <- "naowinter"
amowintervar <- "amowinter"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
          asclayer$o2_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascvar == naowintervar){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascvar == amowintervar){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist2015.csv", row.names = TRUE)
}
```