### 2019-09-24 src/plots for paper.Rmd
plotting for paper - error with names for depths

---
title: "plots for paper"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.

Built with 'r getRversion()'..

# Package dependencies
You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). Note this function checks to see if the packages are installed first before loading.

```{r pre-install & load packages, include=FALSE}
packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster")
source("../src/ipak.R")
ipak(packages)
```

# stacked permutation importance plot and jacknife plots

```{r create permutation & jack data}
maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
perm <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
colnames(perm) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
jack <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
colnames(jack) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")

for (m in 1:length(maxr_lst)){
  maxr <- read.csv(maxr_lst[m])
  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
  perm$Month[m] <- mth
  jack$Month[m] <- mth
  maxt <- tail(maxr, 1)
  perm$amo_sample[m] <- maxt$amo_sample.permutation.importance
  perm$amo_winter[m] <- maxt$amo_winter.permutation.importance
  perm$chl_surface[m] <- maxt$chl_surface.permutation.importance
  perm$nao_prev[m] <- maxt$nao_prev.permutation.importance
  perm$nao_sample[m] <- maxt$nao_sample.permutation.importance
  perm$nao_winter[m] <- maxt$nao_winter.permutation.importance
  perm$o2_depth[m] <- maxt$o2_depth.permutation.importance
  perm$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
  perm$temp_depth[m] <- maxt$temp_depth.permutation.importance

  jack$amo_sample[m] <- maxt$Training.gain.without.amo_sample
  jack$amo_winter[m] <- maxt$Training.gain.without.amo_winter
  jack$chl_surface[m] <- maxt$Training.gain.without.chl_surface
  jack$nao_prev[m] <- maxt$Training.gain.without.nao_prev
  jack$nao_sample[m] <- maxt$Training.gain.without.nao_sample
  jack$nao_winter[m] <- maxt$Training.gain.without.nao_winter
  jack$o2_depth[m] <- maxt$Training.gain.without.o2_depth
  jack$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
  jack$temp_depth[m] <- maxt$Training.gain.without.temp_depth
}

mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
permu <- perm[match(mthord, perm$Month), ] #reorder based on mthord
jackk <- jack[match(mthord, jack$Month), ] #reorder based on mthord

#rename the months so they are capitalised
permu$Month <- capitalize(permu$Month)
jackk$Month <- capitalize(jackk$Month)

#rename the col names so they are capitalised and shortened
colnames(permu)[colnames(permu)=="amo_sample"] <- "AMO_Sample"
colnames(permu)[colnames(permu)=="amo_winter"] <- "AMO_Winter"
colnames(permu)[colnames(permu)=="chl_surface"] <- "Chl_Surface"
colnames(permu)[colnames(permu)=="nao_prev"] <- "NAO_Prev"
colnames(permu)[colnames(permu)=="nao_sample"] <- "NAO_Sample"
colnames(permu)[colnames(permu)=="nao_winter"] <- "NAO_Winter"
colnames(permu)[colnames(permu)=="o2_depth"] <- "O2_Depth"
colnames(permu)[colnames(permu)=="salinity_depth"] <- "Salinity_Depth"
colnames(permu)[colnames(permu)=="temp_depth"] <- "Temp_Depth"

colnames(jackk)[colnames(jackk)=="amo_sample"] <- "AMO_Sample"
colnames(jackk)[colnames(jackk)=="amo_winter"] <- "AMO_Winter"
colnames(jackk)[colnames(jackk)=="chl_surface"] <- "Chl_Surface"
colnames(jackk)[colnames(jackk)=="nao_prev"] <- "NAO_Prev"
colnames(jackk)[colnames(jackk)=="nao_sample"] <- "NAO_Sample"
colnames(jackk)[colnames(jackk)=="nao_winter"] <- "NAO_Winter"
colnames(jackk)[colnames(jackk)=="o2_depth"] <- "O2_Depth"
colnames(jackk)[colnames(jackk)=="salinity_depth"] <- "Salinity_Depth"
colnames(jackk)[colnames(jackk)=="temp_depth"] <- "Temp_Depth"

write.csv(permu, "../output/maxent/models/permutation_avg.csv", row.names = FALSE)
write.csv(jackk, "../output/maxent/models/jack_avg.csv", row.names = FALSE)

```

```{r create jacknife plots}
maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
jack_df <- list()

for (m in 1:length(maxr_lst)){
  res <- read.csv(maxr_lst[m], header = TRUE)
  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
  mth <- capitalize(mth)
  res_df <- tail(res, 1)
  var_df <- res_df[,c(grep("Training.gain.with*",names(res_df)))]
  var_df$id <- 1:nrow(var_df)
  var_df <- melt(var_df, id.vars=c("id"))
  jack_with <- var_df[grep("with.only",var_df$variable),]
  jack_with$Type_of_Run <- "With only the variable"
  jack_without <- var_df[grep("without",var_df$variable),]
  jack_without$Type_of_Run <- "Without the variable"
  gain_all <- res_df[c(grep("Regularized.",names(res_df)))]
  gain_all$id <- 1:nrow(gain_all)
  gain_all <- melt(gain_all, id.vars=c("id"))
  gain_all$Type_of_Run <- "With all variables"
  gain_all$variable <- "all_variables"
  jack_table <- rbind(jack_with,jack_without,gain_all)
  jack_table$state <- mth

  jack_table$variable <- gsub(".*[.]","",jack_table$variable)
  jack_table$variable <- gsub("amo_sample", "AMO_Sample", jack_table$variable)
  jack_table$variable <- gsub("amo_winter", "AMO_Winter", jack_table$variable)
  jack_table$variable <- gsub("chl_surface", "Chl_Surface", jack_table$variable)
  jack_table$variable <- gsub("nao_prev", "NAO_prev", jack_table$variable)
  jack_table$variable <- gsub("nao_sample", "NAO_Sample", jack_table$variable)
  jack_table$variable <- gsub("nao_winter", "NAO_Winter", jack_table$variable)
  jack_table$variable <- gsub("o2_depth", "O2_Depth", jack_table$variable)
  jack_table$variable <- gsub("salinity_depth", "Salinity_Depth", jack_table$variable)
  jack_table$variable <- gsub("temp_depth", "Temp_Depth", jack_table$variable)

  jack_df[[m]] <- jack_table

}

var_jack <- do.call(rbind, jack_df)

var_jack$state <- factor(var_jack$state,levels=c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
var_jack$Type_of_Run <- factor(var_jack$Type_of_Run,levels=c("With all variables", "Without the variable","With only the variable"))
a <- subset(var_jack, state == "Mar" | state == "Apr" |state == "May" | state == "Jun" | state == "Jul")
b <- subset(var_jack, state == "Aug" | state == "Sep" |state == "Oct" | state == "Nov" | state == "Dec")

plotlst <- list(a, b)

color_type_1 <- c("#081d58", "#999999","#1d91c0")
varpos <- c("Temp_Depth", "Salinity_Depth", "O2_Depth", "Chl_Surface", "NAO_Sample", "NAO_Prev", "NAO_Winter", "AMO_Sample", "AMO_Winter")

for (p in 1:length(plotlst)){
  pt <- p
  dev.new()
  jack <- ggplot(plotlst[[p]], aes(variable,value))+
  geom_bar(aes(fill=Type_of_Run),position="dodge",stat="identity")+ coord_flip()+ facet_grid(state~., scales="free", space = "free")+
  theme_bw(14)+
  scale_fill_manual(values=color_type_1,guide=guide_legend(reverse=FALSE))+
  theme(axis.text=element_text(size=8)) +
  theme(legend.title=element_blank())+
  theme(legend.position="top")+
  theme(strip.text=element_text(size=12))+
  theme(strip.text.y = element_text(angle = 0)) +
  labs(x="Variable",y="Jackknife of Regularized Training Gain")
  ggsave(filename = paste0("../output/maxent/paper_plots/jackknife_plot_", pt, ".png"), plot = jack, dpi = 900, height = 9, width = 7)
  dev.off()
}

```

```{r permulation plot}
permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = -1)
png("../output/maxent/paper_plots/permutation_plot.png")
print(perpl)
dev.off()
```

# plot median probability by depth

```{r get median probability for each depth}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

meddf <- data.frame(matrix(ncol = 11, nrow = 49))
colnames(meddf) <- c("depth", mthord)

for (m in 1:length(mthord)){
  mth <- mthord[m]
  mthlst <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
  rstk <- stack(mthlst)
  rsum <- summary(rstk, na.rm = TRUE)
  rsum <- rsum["Median", ]
  meddf[mth] <- rsum
  dep <- rstk@names
  write.csv(meddf, paste0("../output/maxent/paper_plots/med_data/median.csv"), row.names = TRUE)
 # write.csv(rsum, paste0("../output/maxent/paper_plots/med_data/median_", mth, ".csv"), row.names = TRUE)
}

medlst <- list.files(paste0("../output/maxent/paper_plots/med_data"), pattern = "^median.*.csv", full.names = TRUE)
tables <- lapply(medlst, read.csv, header = TRUE)
allmed <- do.call(cbind, tables)

```