### 2019-10-01 src/plots for paper.Rmd
i hate R

---
title: "plots for paper"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.

Built with 'r getRversion()'..

# Package dependencies
You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). Note this function checks to see if the packages are installed first before loading.

```{r pre-install & load packages, include=FALSE}
packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster")
source("../src/ipak.R")
ipak(packages)
```

# stacked permutation importance plot and jacknife plots

```{r create permutation & jack data}
maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
perm <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
colnames(perm) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
jack <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
colnames(jack) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")

for (m in 1:length(maxr_lst)){
  maxr <- read.csv(maxr_lst[m])
  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
  perm$Month[m] <- mth
  jack$Month[m] <- mth
  maxt <- tail(maxr, 1)
  perm$amo_sample[m] <- maxt$amo_sample.permutation.importance
  perm$amo_winter[m] <- maxt$amo_winter.permutation.importance
  perm$chl_surface[m] <- maxt$chl_surface.permutation.importance
  perm$nao_prev[m] <- maxt$nao_prev.permutation.importance
  perm$nao_sample[m] <- maxt$nao_sample.permutation.importance
  perm$nao_winter[m] <- maxt$nao_winter.permutation.importance
  perm$o2_depth[m] <- maxt$o2_depth.permutation.importance
  perm$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
  perm$temp_depth[m] <- maxt$temp_depth.permutation.importance

  jack$amo_sample[m] <- maxt$Training.gain.without.amo_sample
  jack$amo_winter[m] <- maxt$Training.gain.without.amo_winter
  jack$chl_surface[m] <- maxt$Training.gain.without.chl_surface
  jack$nao_prev[m] <- maxt$Training.gain.without.nao_prev
  jack$nao_sample[m] <- maxt$Training.gain.without.nao_sample
  jack$nao_winter[m] <- maxt$Training.gain.without.nao_winter
  jack$o2_depth[m] <- maxt$Training.gain.without.o2_depth
  jack$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
  jack$temp_depth[m] <- maxt$Training.gain.without.temp_depth
}

mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
permu <- perm[match(mthord, perm$Month), ] #reorder based on mthord
jackk <- jack[match(mthord, jack$Month), ] #reorder based on mthord

#rename the months so they are capitalised
permu$Month <- capitalize(permu$Month)
jackk$Month <- capitalize(jackk$Month)

#rename the col names so they are capitalised and shortened
colnames(permu)[colnames(permu)=="amo_sample"] <- "AMO_Sample"
colnames(permu)[colnames(permu)=="amo_winter"] <- "AMO_Winter"
colnames(permu)[colnames(permu)=="chl_surface"] <- "Chl_Surface"
colnames(permu)[colnames(permu)=="nao_prev"] <- "NAO_Prev"
colnames(permu)[colnames(permu)=="nao_sample"] <- "NAO_Sample"
colnames(permu)[colnames(permu)=="nao_winter"] <- "NAO_Winter"
colnames(permu)[colnames(permu)=="o2_depth"] <- "O2_Depth"
colnames(permu)[colnames(permu)=="salinity_depth"] <- "Salinity_Depth"
colnames(permu)[colnames(permu)=="temp_depth"] <- "Temp_Depth"

colnames(jackk)[colnames(jackk)=="amo_sample"] <- "AMO_Sample"
colnames(jackk)[colnames(jackk)=="amo_winter"] <- "AMO_Winter"
colnames(jackk)[colnames(jackk)=="chl_surface"] <- "Chl_Surface"
colnames(jackk)[colnames(jackk)=="nao_prev"] <- "NAO_Prev"
colnames(jackk)[colnames(jackk)=="nao_sample"] <- "NAO_Sample"
colnames(jackk)[colnames(jackk)=="nao_winter"] <- "NAO_Winter"
colnames(jackk)[colnames(jackk)=="o2_depth"] <- "O2_Depth"
colnames(jackk)[colnames(jackk)=="salinity_depth"] <- "Salinity_Depth"
colnames(jackk)[colnames(jackk)=="temp_depth"] <- "Temp_Depth"

write.csv(permu, "../output/maxent/models/permutation_avg.csv", row.names = FALSE)
write.csv(jackk, "../output/maxent/models/jack_avg.csv", row.names = FALSE)

```

```{r create jacknife plots}
maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
jack_df <- list()

for (m in 1:length(maxr_lst)){
  res <- read.csv(maxr_lst[m], header = TRUE)
  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
  mth <- capitalize(mth)
  res_df <- tail(res, 1)
  var_df <- res_df[,c(grep("Training.gain.with*",names(res_df)))]
  var_df$id <- 1:nrow(var_df)
  var_df <- melt(var_df, id.vars=c("id"))
  jack_with <- var_df[grep("with.only",var_df$variable),]
  jack_with$Type_of_Run <- "With only the variable"
  jack_without <- var_df[grep("without",var_df$variable),]
  jack_without$Type_of_Run <- "Without the variable"
  gain_all <- res_df[c(grep("Regularized.",names(res_df)))]
  gain_all$id <- 1:nrow(gain_all)
  gain_all <- melt(gain_all, id.vars=c("id"))
  gain_all$Type_of_Run <- "With all variables"
  gain_all$variable <- "all_variables"
  jack_table <- rbind(jack_with,jack_without,gain_all)
  jack_table$state <- mth

  jack_table$variable <- gsub(".*[.]","",jack_table$variable)
  jack_table$variable <- gsub("amo_sample", "AMO_Sample", jack_table$variable)
  jack_table$variable <- gsub("amo_winter", "AMO_Winter", jack_table$variable)
  jack_table$variable <- gsub("chl_surface", "Chl_Surface", jack_table$variable)
  jack_table$variable <- gsub("nao_prev", "NAO_prev", jack_table$variable)
  jack_table$variable <- gsub("nao_sample", "NAO_Sample", jack_table$variable)
  jack_table$variable <- gsub("nao_winter", "NAO_Winter", jack_table$variable)
  jack_table$variable <- gsub("o2_depth", "O2_Depth", jack_table$variable)
  jack_table$variable <- gsub("salinity_depth", "Salinity_Depth", jack_table$variable)
  jack_table$variable <- gsub("temp_depth", "Temp_Depth", jack_table$variable)

  jack_df[[m]] <- jack_table

}

var_jack <- do.call(rbind, jack_df)

var_jack$state <- factor(var_jack$state,levels=c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
var_jack$Type_of_Run <- factor(var_jack$Type_of_Run,levels=c("With all variables", "Without the variable","With only the variable"))
a <- subset(var_jack, state == "Mar" | state == "Apr" |state == "May" | state == "Jun" | state == "Jul")
b <- subset(var_jack, state == "Aug" | state == "Sep" |state == "Oct" | state == "Nov" | state == "Dec")

plotlst <- list(a, b)

color_type_1 <- c("#081d58", "#999999","#1d91c0")
varpos <- c("Temp_Depth", "Salinity_Depth", "O2_Depth", "Chl_Surface", "NAO_Sample", "NAO_Prev", "NAO_Winter", "AMO_Sample", "AMO_Winter")

for (p in 1:length(plotlst)){
  pt <- p
  dev.new()
  jack <- ggplot(plotlst[[p]], aes(variable,value))+
  geom_bar(aes(fill=Type_of_Run),position="dodge",stat="identity")+ coord_flip()+ facet_grid(state~., scales="free", space = "free")+
  theme_bw(14)+
  scale_fill_manual(values=color_type_1,guide=guide_legend(reverse=FALSE))+
  theme(axis.text=element_text(size=8)) +
  theme(legend.title=element_blank())+
  theme(legend.position="top")+
  theme(strip.text=element_text(size=12))+
  theme(strip.text.y = element_text(angle = 0)) +
  labs(x="Variable",y="Jackknife of Regularized Training Gain")
  ggsave(filename = paste0("../output/maxent/paper_plots/jackknife_plot_", pt, ".png"), plot = jack, dpi = 900, height = 9, width = 7)
  dev.off()
}

```

```{r permulation plot}
permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = -1)
png("../output/maxent/paper_plots/permutation_plot.png")
print(perpl)
dev.off()
```

# plot median probability by depth

```{r get median probability for each depth}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

meddf <- data.frame(matrix(ncol = 3, nrow = 490))
colnames(meddf) <- c("layer", "month", "median")
r <- 1

for (m in 1:length(mthord)){
  mth <- mthord[m]
  mthlst <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
  rstk <- stack(mthlst)
  rsum <- summary(rstk, na.rm = TRUE)
  rsum <- rsum["Median", ]
  for (l in 1:length(rsum)){
    meddf$median[r] <- rsum[l]
    meddf$month[r] <- mth
    meddf$layer[r] <- names(rstk[[l]])
    r <- r+1
  }
}

#rename the layers to be just the depth

meddf$layer <- gsub("avg_cloglog_", "", x = meddf$layer)
meddf <- meddf[order(meddf$layer), ]

write.csv(meddf, paste0("../output/maxent/paper_plots/med_data/median.csv"), row.names = FALSE)
```

```{r}
a <- ggplot(data = meddf) + geom_point(aes(y = layer, x = median, colour = month))
a

```

```{r get max probability for each depth}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

meddf <- data.frame(matrix(ncol = 3, nrow = 490))
colnames(meddf) <- c("layer", "month", "max")
r <- 1

for (m in 1:length(mthord)){
  mth <- mthord[m]
  mthlst <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
  rstk <- stack(mthlst)
  rsum <- summary(rstk, na.rm = TRUE)
  rsum <- rsum["Max.", ]
  for (l in 1:length(rsum)){
    meddf$max[r] <- rsum[l]
    meddf$month[r] <- mth
    meddf$layer[r] <- names(rstk[[l]])
    r <- r+1
  }
}

#rename the layers to be just the depth

meddf$layer <- gsub("avg_cloglog_", "", x = meddf$layer)
meddf$layer <- as.numeric(meddf$layer)
meddf <- meddf[order(meddf$layer), ]

write.csv(meddf, paste0("../output/maxent/paper_plots/med_data/max.csv"), row.names = FALSE)
```

```{r}
a <- ggplot(data = meddf) + geom_point(aes(y = layer, x = max, colour = month))
a
```

#Histograms of probability by depth

count number of cells that are not NA in each raster layer that are inside the mcp

```{r}
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis

clogasc_list <- list.files(paste0("../output/maxent/predictions/nov"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
cumst <- data.frame(0, 0)
colnames(cumst) <- c("depth", "no_cells")
for (cum in 1:length(clogasc_list)){
  cumasc <- raster(clogasc_list[cum])
  ascname <- names(cumasc)
  cumst[cum, 1] <- str_sub(ascname, 13)
  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
  ccnt <- freq(cumasc, useNA='no')
  ccnt <- sum(ccnt[ ,2])
  cumst[cum, 2] <- ccnt
}

cumst$depth <- as.numeric(cumst$depth)
cumst1 <- cumst[order(cumst$no_cells), ]

write.csv(cumst1, "../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", row.names = FALSE)

```

first bin the probabilities for cells inside the mcp
```{r bin probabilities}
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mth <- mthord[m]
  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)

  binprob <- function(x){
    ifelse(x > 0 & x < 0.1, 0.1,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_1_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_1_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.1 & x < 0.2, 0.2,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_2_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_2_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.2 & x < 0.3, 0.3,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_3_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_3_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.3 & x < 0.4, 0.4,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_4_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_4_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.4 & x < 0.5, 0.5,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_5_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_5_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.5 & x < 0.6, 0.6,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_6_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_6_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.6 & x < 0.7, 0.7,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_7_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_7_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.7 & x < 0.8, 0.8,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_8_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_8_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.8 & x < 0.9, 0.9,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_9_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_9_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.9 & x <= 1, 1,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_10_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_10_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }
}
```

get data to plot probability by depth... need to do manually 1-10
```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") 
binlst <- c(1:10)

binprob <- function(x){
  ifelse(x >0,  1,
  ifelse(x == -9999, 0, NA))
}

for (m in 1:length(mthord)){
  for (b in 1:length(binlst)){
  mth <- mthord[m]
  bin <- binlst[b]
  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_binned_", bin, "_.*asc"), full.names = TRUE)

  binaryavgstack <- stack(ascdepth)
  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
  stackstathigh[ ,1] <- c(1:49)
  colnames(stackstathigh) <- c("asclayer", "prob_bin")

  for (lyr in 1:nlayers(binaryavgstack)){
    ascname <-  str_sub(ascdepth[lyr], 63, -5)
    dlay <- binaryavgstack[[lyr]]
    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
    lyrsum <- cellStats(dlay, sum)
    stackstathigh$prob_bin[lyr] <- lyrsum
    stackstathigh$asclayer[lyr] <- ascname

  }
  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_probbin_", bin, "_cellcount_bydepth_avg.csv"), row.names = FALSE)
  }
}

```

```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mth <- mthord[m]
  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, ".*probbin_.*.csv"), full.names = TRUE)
  a <- read.csv(probblst[1])
  b <- read.csv(probblst[2])
  c <- read.csv(probblst[3])
  d <- read.csv(probblst[4])
  e <- read.csv(probblst[5])
  f <- read.csv(probblst[6])
  g <- read.csv(probblst[7])
  h <- read.csv(probblst[8])
  i <- read.csv(probblst[9])
  j <- read.csv(probblst[10])

  stackstat <- cbind(a, b[ ,2], c[ ,2], d[ ,2], e[ , 2], f[ ,2], g[ ,2], h[ ,2], i[ ,2], j[ ,2])
  colnames(stackstat) <- c("depth", "0 - 0.09", "0.1 - 0.19", "0.2 - 0.29", "0.3 - 0.39", "0.4 - 0.49", "0.5 - 0.59", "0.6 - 0.69", "0.7 - 0.79", "0.8 - 0.89", ">0.9")
  stackstat <- melt(stackstat, id = "depth")
  stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]

  colnames(stackstat) <- c("Depth", "Probability", "Cells")

  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
}

```

think about how to get proportions...

```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)

for (m in 1:length(mthord)){
  mth <- mthord[m]
  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
  for (r in 1:nrow(a)){
    a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
    d1 <- str_sub(a$Depth[r], end = 4)
	  for (ro in 1:nrow(cumst1)){
      d2 <- str_sub(cumst1$depth[ro], end = 4)
      if (d1 == d2){
          a$total_cells[r] <- cumst1$no_cells[ro]
      }
	  }
    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
  }
    a$Depth <- as.numeric(a$Depth)

  write.csv(a, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), row.names = FALSE)
}

```

histogram of probability by depth - by proportion

```{r}

bydepth <- ggplot(a, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() #XXXXXcareful - the deeper layers aren't being displayed correctly here
plot(bydepth)
#png(paste0("../output/maxent/predictions/", mthna, "/depthplot_probavg_prop.png"))
#print(bydepth)
#dev.off()
```

ok lets try colour the probabilites to match the maps...

```{r}
probcol <- rev(brewer.pal(10,"RdYlGn"))
bydepth <- ggplot(a, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() + scale_fill_manual(values = probcol) #XXXXXcareful - the deeper layers aren't being displayed correctly here
plot(bydepth)
```

ok now lets lump all the 0.39 and under together...

```{r}
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

binprob <- function(x){
    ifelse(x > 0 & x <0.4, 1,
    ifelse(x <= 0, 0, NA))
}

for (m in 1:length(mthord)){
  mth <- "dec"
  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- raster::mask(x = cumasc, mask = outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_binned_avgprob_", ascname, ".asc"), overwrite = TRUE)
    #plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_binned_avgprob_", ascname, ".png")), overwrite = TRUE)
    #dev.off() # stops automatic saving of the plot to a png
  }
}
```

get data to plot probability by depth... 
```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mth <- mthord[m]
  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_u0.4_binned_avgprob_", ".*asc"), full.names = TRUE)

  binaryavgstack <- stack(ascdepth)
  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
  stackstathigh[ ,1] <- c(1:49)
  colnames(stackstathigh) <- c("asclayer", "prob_bin")

  for (lyr in 1:nlayers(binaryavgstack)){
    ascname <-  str_sub(ascdepth[lyr], 63, -5)
    dlay <- binaryavgstack[[lyr]]
    lyrsum <- cellStats(dlay, sum)
    stackstathigh$prob_bin[lyr] <- lyrsum
    stackstathigh$asclayer[lyr] <- ascname

  }
  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_probbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
}

```

```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mth <- mthord[m]
  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_u0.4_probbin_cellcount_bydepth_avg.csv"), full.names = TRUE)
  stackstat <- read.csv(probblst[1])

  colnames(stackstat) <- c("depth", "0 - 0.39")
  stackstat <- melt(stackstat, id = "depth")
  stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]

  colnames(stackstat) <- c("Depth", "Probability", "Cells")

  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
}

```

get proportions...

```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)

for (m in 1:length(mthord)){
  mth <- mthord[m]
  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
  for (r in 1:nrow(a)){
    a$Depth[r] <- gsub("prob_avg_cloglog_", "", x = a$Depth[r])
    d1 <- str_sub(a$Depth[r], end = 4)
	  for (ro in 1:nrow(cumst1)){
      d2 <- str_sub(cumst1$depth[ro], end = 4)
      if (d1 == d2){
          a$total_cells[r] <- cumst1$no_cells[ro]
      }
	  }
    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
  }
   a$Depth <- as.numeric(a$Depth)

  write.csv(a, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_dis_prob_cells.csv"), row.names = FALSE)
}

```

merge the u0.4 dataset with the original
```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
for (m in 1:length(mthord)){
  mth <- mthord[m]
  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
  b <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
  c <- rbind(a, b)
  c <- c[!(c$Probability == "0 - 0.09"), ]
  c <- c[!(c$Probability == "0.1 - 0.19"), ]
  c <- c[!(c$Probability == "0.2 - 0.29"), ]
  c <- c[!(c$Probability == "0.3 - 0.39"), ]
  write.csv(c, paste0("../output//maxent/paper_plots/prob_by_depth/", mth, "_u0.4plus_dis_prob_cells.csv"))
}
```

ok no try plot

```{r}
probcol <- rev(brewer.pal(7,"RdYlGn"))
bydepth <- ggplot(c, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() + scale_fill_manual(values = probcol) #XXXXXcareful - the deeper layers aren't being displayed correctly here
plot(bydepth)
```

```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)
for (m in 1:length(mthord)){
  mth <- mthord[m]
  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), header = TRUE)
  head(a)
  d <- unique(a$Depth)
  x <- data.frame(matrix(nrow = length(d)))
  for (s in 1:length(d)){
    dep <- d[s]
    sd <- subset(a, Depth == dep)
    sd <- subset(sd, Probability == "0 - 0.09"| Probability == "0.1 - 0.19"| Probability == "0.2 - 0.29"| Probability == "0.3 - 0.39")
    c <- sum(sd$Cells)
    x$Depth[s] <- dep
    x$Probability[s] <- "0 - 0.39"
    x$cells[s] <- c
    x$p_cells[s] <- 1
  }
  x <- subset(x, select = -c(matrix.nrow...length.d..))
  x$cells[is.na(x$cells)] <- 0
  # for (r in 1:nrow(x)){
  #   d1 <- str_sub(x$Depth[r], end = 4)
  #   for (ro in 1:nrow(cumst1)){
  #     d2 <- str_sub(cumst1$depth[ro], end = 4)
  #     if (d1 == d2){
  #         x$total_cells[r] <- cumst1$no_cells[ro]
  #     }
  #   }
  # }
  # for (r in 1:nrow(x)){
  #    x$p_cells[r] <- (x$Cells[r]/x$total_cells[r])*100

 # }

  write.csv(x, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4plus_dis_prob_cells.csv"))
}
```

merge the u0.4 dataset with the original
```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
for (m in 1:length(mthord)){
  mth <- mthord[m]
  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4plus_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
  b <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
  c <- rbind(a, b)
  c <- c[!(c$Probability == "0 - 0.09"), ]
  c <- c[!(c$Probability == "0.1 - 0.19"), ]
  c <- c[!(c$Probability == "0.2 - 0.29"), ]
  c <- c[!(c$Probability == "0.3 - 0.39"), ]
  write.csv(c, paste0("../output//maxent/paper_plots/prob_by_depth/", mth, "_u0.4s_dis_prob_cells.csv"))
}
```

```{r}
probcol <- rev(brewer.pal(7,"RdYlGn"))
bydepth <- ggplot(c, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() + scale_fill_manual(values = probcol) #XXXXXcareful - the deeper layers aren't being displayed correctly here
plot(bydepth)
```

ok dont forget the colours will need to match the maps so you may need to manually create your own list for whatever scheme you choose...

now MJ suggested to only show some depths (the most commonly predicted ones)

ok lets try get an probability for the whole month

```{r}
a <- read.csv("../output/maxent/predictions/jul/dis_prob_cells.csv", header = TRUE)

#subset the probability ranges
p1 <- subset(a, Probability == "0 - 0.09")
p2 <- subset(a, Probability == "0.1 - 0.19")
p3 <- subset(a, Probability == "0.2 - 0.29")
p4 <- subset(a, Probability == "0.3 - 0.39")
p5 <- subset(a, Probability == "0.4 - 0.49")
p6 <- subset(a, Probability == "0.5 - 0.59")
p7 <- subset(a, Probability == "0.6 - 0.69")
p8 <- subset(a, Probability == "0.7 - 0.79")
p9 <- subset(a, Probability == "0.8 - 0.89")
p10 <- subset(a, Probability == ">0.9")

#count the number of cells per probability.

cp1 <- sum(p1$Cells)
cp2 <- sum(p2$Cells)
cp3 <- sum(p3$Cells)
cp4 <- sum(p4$Cells)
cp5 <- sum(p5$Cells)
cp6 <- sum(p6$Cells)
cp7 <- sum(p7$Cells)
cp8 <- sum(p8$Cells)
cp9 <- sum(p9$Cells)
cp10 <- sum(p10$Cells)

tc2 <- sum(cp1, cp2, cp3, cp4, cp5, cp6, cp7, cp8, cp9, cp10) #how many cells in total for novy have a value
tcell <- sum(p1$total_cells) # how many cells are available to have a value...
#tc2 and tcell should match

#create datbase of total cells per probability bin for the month

b <- unique(a$Probability) #get the unique probability bins
c <- as.data.frame(rbind(cp1, cp2,  cp3, cp4, cp5, cp6, cp7, cp8, cp9, cp10))
c$probability <- b
colnames(c) <- c("cells", "prob_bin")
rownames(c) <- c()
c$month <- "jul"
```

make plot of probability over the month
```{r}
pmth <- ggplot(d, aes(month, cells, fill = prob_bin)) + geom_bar(stat='identity') 
plot(pmth)
png(paste0("../output/maxent/predictions/", mthna, "/prob_avg_mth.png"))
print(pmth)
dev.off()
```

merge two months of data together into one file then plot

```{r}
f <- rbind(c, d)
pmth <- ggplot(f, aes(month, cells, fill = prob_bin)) + geom_bar(stat='identity') 
plot(pmth)
png(paste0("../output/maxent/predictions/prob_avg_mth_test.png"))
print(pmth)
dev.off()
```