### 2020-02-12 src/comb-maxent.Rmd
updated plots (background point gen)

  
```{r pre-install & load packages, include=FALSE}
options(java.parameters = "-Xmx2g" )#needs to run before dismo loaded - boosts memory for java to 1g
packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr", "reshape2", "tidyverse", "patchwork", "purrr")
source("../src/ipak.R")
ipak(packages)
```

#need to reclassify cells outside the mcp to NA for predictions
```{r}
outsidemcp <- shapefile("../data/env/landp/outside-mcp-wgs84.shp") #made in arcgis - in aea

ascdirs <- list.dirs("../data/env/ascii-original", full.names = FALSE, recursive = FALSE)

for (ad in 1:length(ascdirs)){
  timeslice <- ascdirs[ad]
  ascfiles <- list.files(paste0("../data/env/ascii-original/", timeslice), pattern = "\\.asc$", full.names = TRUE)
  for (file in 1:length(ascfiles)){
    ras<- raster(ascfiles[file])
    crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
    ascname <- str_sub(ascfiles[file], start = 38)
    ras[outsidemcp] <- 500000
    ras[ras == 500000] <- NA
    writeRaster(ras, paste0("../data/env/asciitrim/", ascdirs[ad], "/", ascname), overwrite=TRUE)
  }
}

```

load the presence-background dataset 
("../output/bio/presab_nafo_yr.csv")

select month
```{r month of interest -change}
mthna <- "dec"
mthno <- "12"
```

```{r load pres-background dataset}
presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
colnames(presback)
```

remove all rows with the year 2015 (due to low number of samples)
```{r}
presback <- subset(presback, year!="2015")
```

this dataset contains a lot of columns you don't need for the maxent models. Remove them here

```{r remove unnecessary columns}
presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt))
colnames(presback)
```

monthly model data selection and prep
For each monthly model, subset the data to the month you want, and the variables you want (Based on spearmans and VIF)

```{r}
marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
aprvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
augvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
decvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")

mthnavar_lst <- list(mar = marvar, apr = aprvar, may = mayvar, jun = junvar, jul = julvar, aug = augvar, sep = sepvar, oct = octvar, nov = novvar, dec = decvar)
```

```{r subsest data by month}
prback <- subset(presback, month == mthno)
prback <- subset(prback, select = mthnavar_lst[[mthna]])
```

now each month contains NAs - lets remove these (maxent principle doesnt allow for missing data - http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.15.5607&rep=rep1&type=pdf)

note will need to reajust background points too as some years may no longer have presence data

```{r remove na vals}
prback <- na.omit(prback)
presyrs <-  subset(prback, occurrence == 1) #this is to get years with complete presence data 
presyrs <- unique(presyrs$year) #make a vector of unique years
prback <- prback[prback$year %in% presyrs, ] #keep only the data for which there is precence years

```

```{r final models plus predict}
modelmonth <- cbind(paste0(mthna, unique(prback$year)))
aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
unlayer <- read.csv(paste0("../data/env/test-comb/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr

 for (year in 1:nrow(modelmonth)){
   timeslice <- modelmonth[year] #testing timeslice
   yr <- str_sub(timeslice, -4) #the year being used for testing
   train <- subset(prback, year != yr) #training data is all years but testing
   trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
   test <-  subset(prback, year == yr) #creates the testing dataset
   testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
   testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
   testb <- subset(test, occurrence == "0") #splits the testing dataset into background
   testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
   #mess_pr <- subset(train, occurrence == 1)
   #mess_pr  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
   train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

   if (yr == "1998"){
     betaval <- 0.25
   } else if (yr == "2001"){
     betaval <- 0.75
   } else if (yr == "2002" | yr == "2003"| yr == "2004"| yr == "2006"| yr == "2011"){
     betaval <- 1
   } else if (yr == "2005"){
     betaval <- 1.25
   } else if (yr == "1999" | yr == "2000"| yr == "2009"){
     betaval <- 1.5
   } else if (yr == "2007"){
     betaval <- 1.75
   } else if (yr == "2010"){
     betaval <- 2
   } else if (yr == "2008"){
     betaval <- 2.25
   }

     mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/test-comb/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))

     saveRDS(mod, file= paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))

     #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code

     #count number of non-zero coefficients, excluding the final four rows
     coefile <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
     coefile <- head(coefile, -4)
     coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
     nocoeff <- nrow(coefile2)
     bick <- sum(as.numeric(coefile2[, 2]))
     bic <- (bick*log(sum(trainpa == 1)) - 2*11)

     #no presences used in model (less those used for testing)
     maxresults <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
     notrainingpoints <- maxresults$X.Training.samples
     testauc <- maxresults$Test.AUC
     trainauc <- maxresults$Training.AUC

     #natural log
     samplepred <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
     samplepred$nat_log <- log(samplepred$Raw.prediction)
     sumsamplepred <- sum(samplepred$nat_log)

     #test auc scores and other lovely things you might want to include
     ev <- evaluate(p = testp, a= testb, model = mod)
     evauc <- ev@auc
     evcor <- ev@cor
     evpcor <- ev@pcor
     evnp <- ev@np
     evna <- ev@na
     evmaxtprtnr <- threshold(ev, 'spec_sens')
     evmaxkap <- threshold(ev, 'kappa')
     evno_omission <- threshold(ev, 'no_omission')
     evprevalence <- threshold(ev, 'prevalence')
     evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
     evsensitivity <- threshold(ev, 'sensitivity')

     #continuous boyce index
     predPres <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
     predBg <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
     predPres <- predPres$Cloglog.prediction
     predBg <- predBg$Cloglog
     source("../src/contBoyce.R")
     cbi <- contBoyce(pres = predPres, bg = predBg)

       aicetc$train_AUC <- trainauc #training only AUC - not tested
       aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
       aicetc$coefficients <- nocoeff #this is to calculate the AIC
       aicetc$sum.of.logs <- sumsamplepred #this is to calculate the AIC
       aicetc$test_AUC <- evauc #testing AUC
       aicetc$test_evcor <- evcor #testing correlation coefficient
       aicetc$test_evpcor <- evpcor # testing p-value for correlation coefficient
       aicetc$test_nopresence <- evnp
       aicetc$test_nobackground <- evna
       aicetc$test_maxsumsenspec <- evmaxtprtnr
       aicetc$test_maxkap <- evmaxkap
       aicetc$test_no_omission <- evno_omission
       aicetc$test_prevalence <- evprevalence
       aicetc$test_equal_sens_spec <- evequal_sens_spec
       aicetc$test_sensitivity <- evsensitivity
       aicetc$tss <- max(ev@TPR + ev@TNR) - 1
       aicetc$test_bic <- bic #not sure this is correct...
       aicetc$beta <- betaval
       aicetc$cbi <- cbi

     aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
     aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$no_trainng_points/(aicetc$no_trainng_points - aicetc$coefficients-1))))

     #calculate minimal predicted area using ecospat
     aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
     aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
     aicetc$year <- yr

     write.csv(aicetc, paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

    mod <- readRDS(paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))

    yr <- as.numeric(yr)
 unlayer_yr <- subset(unlayer, ascyearslist == yr)
    #unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdeplst <- list()
    stackdep <- subset(unlayer_yr, select = c(stackname))
    stackdeplst[[year]] <- stackdep
    write.csv(stackdep, paste0("../output/maxent/models/test-comb/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)

  for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      #temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      #chl_depth <- projectRaster(chl_depth, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      #o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      #salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      #nao_sample <- projectRaster(nao_sample, aea)
      nao_prev <- raster(unlayer_yr$nao_prev[td])
      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_prev) <- "nao_prev"
      #nao_prev <- projectRaster(nao_prev, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      #nao_winter <- projectRaster(nao_winter, aea)
      amo_sample <- raster(unlayer_yr$amo_sample[td])
      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(amo_sample) <- "amo_sample"
      #amo_sample <- projectRaster(amo_sample, aea)
      amo_winter <- raster(unlayer_yr$amo_winter[td])
      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(amo_winter) <- "amo_winter"
      #amo_winter <- projectRaster(amo_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)

      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=cloglog')) 
      # predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".png")))
      dev.off() # stops automatic saving of the plot to a png
  }
}
```

all layers need to be spit in bash + poweshell
- pre-split.ps1 
- separate.sh
- aftersplit.ps1

# average the maxent results

```{r avg maxentReg}
monthlist <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by

for (mth in 1:length(monthlist)){
    mthna <- monthlist[mth]
    modlist <- list.dirs("../output/maxent/models/test-comb", full.names = FALSE, recursive = FALSE) 
    modlist <- modlist[ grepl(mthna, modlist) ]
    aiclst <- list()
    maxrlst <- list()
    for (i in 1:length(modlist)){
      timeslice <- modlist[i]
      yr <- str_sub(timeslice, -4) #the year 
      if(yr == 1998){
        aic1998 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic1998
        maxr1998 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr1998
        } else if (yr == 1999){
        aic1999 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic1999
        maxr1999 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr1999
        } else if (yr == 2000){
        aic2000 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2000
        maxr2000 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2000
        } else if (yr == 2001){
        aic2001 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2001
        maxr2001 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2001
        } else if (yr == 2002){
        aic2002 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2002
        maxr2002 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2002
        } else if (yr == 2003){
        aic2003 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2003
        maxr2003 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2003
        } else if (yr == 2004){
        aic2004 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2004
        maxr2004 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2004
        } else if (yr == 2005){
        aic2005 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2005
        maxr2005 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2005
        } else if (yr == 2006){
        aic2006 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2006
        maxr2006 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2006
        } else if (yr == 2007){
        aic2007 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2007
        maxr2007 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2007
        } else if (yr == 2008){
        aic2008 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2008
        maxr2008 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2008
        } else if (yr == 2009){
        aic2009 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2009
        maxr2009 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2009
        } else if (yr == 2010){
        aic2010 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2010
        maxr2010 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2010
        } else if (yr == 2011){
        aic2011 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2011
        maxr2011 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2011
        } else if (yr == 2012){
        aic2012 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2012
        maxr2012 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2012
        } else if (yr == 2013){
        aic2013 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2013
        maxr2013 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2013
        } else if (yr == 2014){
        aic2014 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2014
        maxr2014 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2014
        }
    }
  aic <- do.call(rbind, aiclst)
  maxr <- do.call(rbind, maxrlst)

  aicm <- colMeans(aic)
  aicm$mod <- "avg"
  avgAIC <- rbind(aic, aicm)

  maxr$Species <- 1
  maxrm <- colMeans(maxr)
  maxrm$Species <- "avg"
  maxrm <- rbind(maxr, maxrm)

  write.csv(avgAIC, paste0("../output/maxent/models/test-comb/", mthna, "avgAIC.csv"), row.names = FALSE)
  write.csv(maxrm, paste0("../output/maxent/models/test-comb/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
}
```

summary table of average maxent results for each month
```{r average AIC etc for each month in one table}
aic_lst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)
aicsum <- data.frame(matrix(ncol = 11, nrow = length(aic_lst)))
colnames(aicsum) <- c("Month", "BetaVal", "AICc", "Test AUC", "Training AUC", "Correlation", "Correlation P-Val", "TSS", "CBI", "No Omissions", "Prevalence")

for (a in 1:length(aic_lst)){
  aicr <- read.csv(aic_lst[a])
  mth <- str_sub(aic_lst[a], start = 35, end = 37) #selects the month from the filename
  aicsum$Month[a] <- mth
  aicres <- tail(aicr, 1)
  aicsum$BetaVal[a] <- aicres$beta
  aicsum$AICc[a] <- aicres$AICc
  aicsum$`Test AUC`[a] <- aicres$test_AUC
  aicsum$`Training AUC`[a] <- aicres$train_AUC
  aicsum$Correlation[a] <- aicres$test_evcor
  aicsum$`Correlation P-Val` <- aicres$test_evpcor
  aicsum$TSS[a] <- aicres$tss
  aicsum$CBI[a] <- aicres$cbi
  aicsum$`No Omissions`[a] <- aicres$test_no_omission
  aicsum$Prevalence[a] <- aicres$test_prevalence
}

mthord <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
aicsum <- aicsum[match(mthord, aicsum$Month), ] #reorder based on mthord
write.csv(aicsum, "../output/maxent/models/test-comb/aicsummary.csv", row.names = FALSE)

```

# average the ascii layers 
need a montly layer for each month-depth

first need to rebuild a list of stackdepths as this does not work in the code above (due to change in ascii formatting - code above is old)

```{r}
#need an list of the month years
p <- list.dirs("../output/maxent/models/test-comb", recursive = FALSE, full.names = FALSE)
yrmthunique <- as.data.frame(p, stringsAsFactors=FALSE)

for(time in 1:nrow(yrmthunique)){
  slice <- yrmthunique$p[time]
  mthna <- str_sub(slice, end = 3)
  yr <-  str_sub(slice, start = 4)
  mthyrclog <- list.files(paste0("../output/maxent/predictions/com-split/", mthna), pattern = yr)
  stackd <- data.frame(mthyrclog)
  for (st in 1:nrow(stackd)){
    stackd$ascdepths[st] <- str_sub(stackd$mthyrclog[st], 17, -5)
  }
  names(stackd)[names(stackd) == "mthyrclog"] <- "stackname"
  write.csv(stackd, paste0("../output/maxent/models/test-comb/", mthna, yr, "/stackdepthlist_", yr, ".csv"))
}
```

```{r create list of layers}
p <- list.dirs("../output/maxent/models/test-comb", recursive = FALSE, full.names = FALSE)
year <- "2006"

monthlist <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by

for (m in 1:length(monthlist)){
  mthna <- monthlist[m]
  timeslice <- paste(mthna, year, sep = '')
  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
  stlst <- list.files(paste0("../output/maxent/models/test-comb/", timeslice), pattern = "stackdepth.*.csv", full.names = TRUE)
  stdep <- read.csv(stlst[1])
  depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
  names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
  depthlayers$depth <- as.character(depthlayers$depth)

  colno <- 1

  for (d in 1:nrow(depthlayers)){
    dep <- depthlayers$depth[d]
    for (a in 1:length(clogasc_list)){
      asc <- clogasc_list[a]
      ascd <- str_sub(asc, 60, -5)
      if (grepl(dep, ascd)){
        depthlayers[d , paste0("a", colno)] <- asc
         colno <- colno+1
      }
    }
  }
  write.csv(depthlayers, (paste0("../output/maxent/predictions/com-split/", mthna, "/cloglog_asclayerbydepth.csv")), row.names = TRUE)
}
```

```{r avg asc layers cloglog}
mthna <- "dec" #change me
depthlayers <- read.csv(paste0("../output/maxent/predictions/com-split/", mthna, "/cloglog_asclayerbydepth.csv"), header = TRUE)
p <- list.dirs("../output/maxent/models/test-comb", recursive = FALSE, full.names = FALSE)
modlist <- p[ grepl(mthna, p) ]
print(length(modlist)) #need this to modify code below
print(modlist) #need this to modify code below

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1) #v1 just a number
  asclay$V2 <- as.character(asclay$V2) #v2 corresponds to depth
  asclay$V3 <- as.character(asclay$V3) #yr 1
  asclay$V4 <- as.character(asclay$V4) #yr 2
  asclay$V5 <- as.character(asclay$V5) #yr 3
  asclay$V6 <- as.character(asclay$V6) #yr 4
  asclay$V7 <- as.character(asclay$V7)
  asclay$V8 <- as.character(asclay$V8) #yr 6
  asclay$V9 <- as.character(asclay$V9)
  asclay$V10 <- as.character(asclay$V10) #yr 8
  asclay$V11 <- as.character(asclay$V11)
  asclay$V12 <- as.character(asclay$V12) #yr 10
  asclay$V13 <- as.character(asclay$V13)
  asclay$V14 <- as.character(asclay$V14) #yr 12
  asclay$V15 <- as.character(asclay$V15)
  asclay$V16 <- as.character(asclay$V16) #yr 14
  #asclay$V17<- as.character(asclay$V17)
  #asclay$V18 <- as.character(asclay$V18) #yr 16
  #asclay$V19 <- as.character(asclay$V19) #yr 17

  asclaydep <- asclay$V2

  asc1 <- raster(asclay$V3) #1
  asc2 <- raster(asclay$V4)
  asc3 <- raster(asclay$V5) #3
  asc4 <- raster(asclay$V6)
  asc5 <- raster(asclay$V7) #5
  asc6 <- raster(asclay$V8)
  asc7 <- raster(asclay$V9) #7
  asc8 <- raster(asclay$V10)
  asc9 <- raster(asclay$V11) #9
  asc10 <- raster(asclay$V12)
  asc11 <- raster(asclay$V13) #11
  asc12 <- raster(asclay$V14)
  asc13 <- raster(asclay$V15) #13
  asc14 <- raster(asclay$V16)
  #asc15 <- raster(asclay$V17) #15
 # asc16 <- raster(asclay$V18)
 # asc17 <- raster(asclay$V19)

  stkasc <- stack(asc1, asc2, asc3, asc4, asc5, asc6, asc7, asc8, asc9, asc10, asc11, asc12, asc13, asc14) #, asc15) #, asc16) #, asc2015)   asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2009

  avg <- mean(stkasc, na.rm = TRUE)

  writeRaster(avg, paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_", asclaydep, ".asc"), overwrite=TRUE)
  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_", asclaydep, ".png"))) 
  dev.off()

file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_000120.asc"), paste0("../output/maxent/predictions/com-split", mthna, "/avg_cloglog_120.asc"))
file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_000120.png"), paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_120.png"))
file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_041.18.asc"), paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_41.18.asc"))
file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_041.18.png"), paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_41.18.png"))
}
```

#binned probability cell count for the whole 3d space per month

```{r}
library(stringr)

#calculate no of cells in the 3D space
prob <- read.csv("../output/maxent/paper_plots/prob_by_depth/apr_dis_prob_cells.csv", header = TRUE) #just a randomly chosen dataset
t2 <- subset(prob, Probability == "0 - 0.09") #could have chosen any probability
tc <- sum(t2$total_cells) #173623

dislst <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "dis_prob_cells.csv", full.names = TRUE)
totlst <- list()
for (m in 1:length(dislst)){
  pr <- read.csv(dislst[m])
  mth <- str_sub(dislst[m], start = 44, end = 46)
  pr2 <- aggregate(pr$Cells, by=list(Category = pr$Probability), FUN = sum)
  pr2$percent <- (pr2$x/tc)*100
  pr2$month <- mth
  write.csv(pr2, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_cellprob_totals.csv"), row.names = FALSE)
  totlst[[m]] <- pr2
}

#then i just made a table manually in excel
```

#plotting

need some additional packages installed
```{r pre-install & load packages, include=FALSE}
packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer", "cowplot", "rasterVis")
source("../src/ipak.R")
ipak(packages)
```

# stacked permutation importance plot

note if you change mind and want jackknife then the code is in the older plots for papers rmd

first need to make the data plot-ready

```{r create permutation & jack data}
maxr_lst <- list.files("../output/maxent/models/test-comb/", pattern = "maxentresults.csv", full.names = TRUE)
perm <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
colnames(perm) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
# jack <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
# colnames(jack) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")

for (m in 1:length(maxr_lst)){
  maxr <- read.csv(maxr_lst[m])
  mth <- str_sub(maxr_lst[m], start = 35, end = 37) #selects the month from the filename
  perm$Month[m] <- mth
  # jack$Month[m] <- mth
  maxt <- tail(maxr, 1)
  perm$amo_sample[m] <- maxt$amo_sample.permutation.importance
  perm$amo_winter[m] <- maxt$amo_winter.permutation.importance
  perm$chl_surface[m] <- maxt$chl_surface.permutation.importance
  perm$nao_prev[m] <- maxt$nao_prev.permutation.importance
  perm$nao_sample[m] <- maxt$nao_sample.permutation.importance
  perm$nao_winter[m] <- maxt$nao_winter.permutation.importance
  perm$o2_depth[m] <- maxt$o2_depth.permutation.importance
  perm$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
  perm$temp_depth[m] <- maxt$temp_depth.permutation.importance

  # jack$amo_sample[m] <- maxt$Training.gain.without.amo_sample
  # jack$amo_winter[m] <- maxt$Training.gain.without.amo_winter
  # jack$chl_surface[m] <- maxt$Training.gain.without.chl_surface
  # jack$nao_prev[m] <- maxt$Training.gain.without.nao_prev
  # jack$nao_sample[m] <- maxt$Training.gain.without.nao_sample
  # jack$nao_winter[m] <- maxt$Training.gain.without.nao_winter
  # jack$o2_depth[m] <- maxt$Training.gain.without.o2_depth
  # jack$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
  # jack$temp_depth[m] <- maxt$Training.gain.without.temp_depth
}

mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
permu <- perm[match(mthord, perm$Month), ] #reorder based on mthord
# jackk <- jack[match(mthord, jack$Month), ] #reorder based on mthord

# rename the months so they are capitalised
permu$Month <- capitalize(permu$Month)
# jackk$Month <- capitalize(jackk$Month)

# rename the col names so they are capitalised and shortened
colnames(permu)[colnames(permu)=="amo_sample"] <- "AMO_Sample"
colnames(permu)[colnames(permu)=="amo_winter"] <- "AMO_Winter"
colnames(permu)[colnames(permu)=="chl_surface"] <- "Chl_Surface"
colnames(permu)[colnames(permu)=="nao_prev"] <- "NAO_Prev"
colnames(permu)[colnames(permu)=="nao_sample"] <- "NAO_Sample"
colnames(permu)[colnames(permu)=="nao_winter"] <- "NAO_Winter"
colnames(permu)[colnames(permu)=="o2_depth"] <- "O2_Depth"
colnames(permu)[colnames(permu)=="salinity_depth"] <- "Salinity_Depth"
colnames(permu)[colnames(permu)=="temp_depth"] <- "Temp_Depth"

# colnames(jackk)[colnames(jackk)=="amo_sample"] <- "AMO_Sample"
# colnames(jackk)[colnames(jackk)=="amo_winter"] <- "AMO_Winter"
# colnames(jackk)[colnames(jackk)=="chl_surface"] <- "Chl_Surface"
# colnames(jackk)[colnames(jackk)=="nao_prev"] <- "NAO_Prev"
# colnames(jackk)[colnames(jackk)=="nao_sample"] <- "NAO_Sample"
# colnames(jackk)[colnames(jackk)=="nao_winter"] <- "NAO_Winter"
# colnames(jackk)[colnames(jackk)=="o2_depth"] <- "O2_Depth"
# colnames(jackk)[colnames(jackk)=="salinity_depth"] <- "Salinity_Depth"
# colnames(jackk)[colnames(jackk)=="temp_depth"] <- "Temp_Depth"

write.csv(permu, "../output/maxent/models/test-comb/permutation_avg.csv", row.names = FALSE)
# write.csv(jackk, "../output/maxent/models/test-comb/jack_avg.csv", row.names = FALSE)

```

now make the plot

```{r permulation plot}
permu <- read.csv("../output/maxent/models/test-comb/permutation_avg.csv")

mthna <- c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
permelt$variable <- factor(permelt$variable, levels = c("AMO_Sample", "AMO_Winter", "NAO_Sample", "NAO_Prev", "NAO_Winter", "Chl_Surface", "O2_Depth", "Salinity_Depth", "Temp_Depth")) #reorder the variables
perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = 1, labels = c("AMO Sample", "AMO Winter", "NAO Sample", "NAO Prev", "NAO Winter", "Chl Surface", "O2 Depth", "Salinity Depth", "Temp Depth")) + theme_half_open(12) + scale_x_discrete(limits = mthna) + labs(y = "Permutation Importance (%)")
png("../output/maxent/paper_plots/permutation_plot.png")
print(perpl)
dev.off()
```

#historgams of probability by depth

count number of cells that are not NA in each raster layer that are inside the mcp

```{r}

clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/mar"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
cumst <- data.frame(0, 0)
colnames(cumst) <- c("depth", "no_cells")
for (cum in 1:length(clogasc_list)){
  cumasc <- raster(clogasc_list[cum])
  ascname <- names(cumasc)
  cumst[cum, 1] <- str_sub(ascname, 13)
  #cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
  ccnt <- freq(cumasc, useNA='no')
  ccnt <- sum(ccnt[ ,2])
  cumst[cum, 2] <- ccnt
}

cumst$depth <- as.numeric(cumst$depth)
cumst1 <- cumst[order(cumst$no_cells), ]

write.csv(cumst1, "../output/maxent/paper_plots/cellsperlayer.csv", row.names = FALSE)

```

histogram of probability by depth - by proportion

#max kappa thresholds

```{r}
binprob <- function(x){
  ifelse(x > 0 & x >= mk, 1,
  ifelse(x <= 0, 0, NA))
}

alst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)

for (f in 1:length(alst)){
  mth <- str_sub(alst[f], start = 35, end = 37)
  af <- read.csv(alst[f], header = TRUE)
  af <- tail(af, n = 1)
  mk <- af$test_maxkap

  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }
}

```

get data to plot probability by depth... 
```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
binlst <- c(1:10)

binprob <- function(x){
  ifelse(x >0,  1,
  ifelse(x == -9999, 0, NA))
}

for (m in 1:length(mthord)){
  for (b in 1:length(binlst)){
    mth <- mthord[m]
    bin <- binlst[b]
    # mth <- "dec"
    # bin <- 1
  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_maxkap.*asc"), full.names = TRUE)

  binaryavgstack <- stack(ascdepth)
  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
  stackstathigh[ ,1] <- c(1:47)
  colnames(stackstathigh) <- c("asclayer", "prob_bin")

  for (lyr in 1:nlayers(binaryavgstack)){
    ascname <-  str_sub(ascdepth[lyr], 77, -5)
    dlay <- binaryavgstack[[lyr]]
    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
    lyrsum <- cellStats(dlay, sum)
    stackstathigh$prob_bin[lyr] <- lyrsum
    stackstathigh$asclayer[lyr] <- ascname

  }
  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", bin, "_cellcount_bydepth_avg.csv"), row.names = FALSE)
  }
}

```

```{r}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mth <- mthord[m]
  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_probbin_.*.csv"), full.names = TRUE)
  a <- read.csv(probblst[1])
  b <- read.csv(probblst[2])
  c <- read.csv(probblst[3])
  d <- read.csv(probblst[4])
  e <- read.csv(probblst[5])
  f <- read.csv(probblst[6])
  g <- read.csv(probblst[7])
  h <- read.csv(probblst[8])
  i <- read.csv(probblst[9])
  j <- read.csv(probblst[10])

  stackstat <- cbind(a, b[ ,2], c[ ,2], d[ ,2], e[ , 2], f[ ,2], g[ ,2], h[ ,2], i[ ,2], j[ ,2])
  colnames(stackstat) <- c("depth", "0 - 0.09", "0.1 - 0.19", "0.2 - 0.29", "0.3 - 0.39", "0.4 - 0.49", "0.5 - 0.59", "0.6 - 0.69", "0.7 - 0.79", "0.8 - 0.89", ">0.9")
  stackstat <- melt(stackstat, id = "depth")
  stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]

  colnames(stackstat) <- c("Depth", "Probability", "Cells")

  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
}

```

think about how to get proportions...

```{r}
mthord <- c("mar",  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
cumst1 <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE)

for (m in 1:length(mthord)){
  mth <- mthord[m]
  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
  for (r in 1:nrow(a)){
    # a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
    # d1 <- str_sub(a$Depth[r], end = 4)
    d1 <- a$Depth[r]
	  for (ro in 1:nrow(cumst1)){
      #d2 <- str_sub(cumst1$depth[ro], end = 4)
	    d2 <- cumst1$depth[ro]
      if (d1 == d2){
          a$total_cells[r] <- cumst1$no_cells[ro]
      }
	  }
    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
  }
    a$Depth <- as.numeric(a$Depth)

  write.csv(a, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), row.names = FALSE)
}

```
based on Remi's code
```{r}
require(data.table)
require(purrr)
require(lubridate)
require(tidyverse)

filenames <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)
filenames <- filenames[ !grepl("maxkap", filenames) ] #remove all the maxkap files

all <- map_dfr(filenames, function(x) fread(x) %>% 
                       mutate(Month=substr(basename(x),1,3))) %>% 
  mutate(Probability = gsub(">0.9", "0.9 - 1", Probability),
         Depth = factor(round(Depth,1)),
         Depth = factor(Depth,ordered = TRUE, levels = rev(levels(Depth))),
         Month=factor(months(parse_date_time(Month,"%m")),
                      ordered = TRUE,
                      levels=months(parse_date_time(1:12,"%m"))))
#create a list of depths so you can set breaks for the x-axis (which is showing as the y-axis...)
dlst <- unique(all$Depth)
dlst <- as.character(dlst)
dlst <- as.numeric(dlst)
dlst <- sort(dlst)
#make the plot
  ggplot(all) +
  geom_col(aes(Depth, p_cells, fill = Probability),width=1) +
  coord_flip()+
  facet_wrap(~Month,ncol=2)+
  labs(y = "% Ocean at Depth", x = "Depth (meters)")+
  scale_fill_viridis(option="magma", discrete = TRUE, direction = -1, guide=guide_legend(nrow=2, byrow = TRUE))+
  theme_classic()+
  theme(legend.position = "bottom", axis.text.y = element_text(size=7), strip.background = element_blank()) +
  scale_x_discrete(breaks = c(dlst[1], dlst[5], dlst[10], dlst[15], dlst[20], dlst[25], dlst[30], dlst[35], dlst[40], dlst[47]))

ggsave("../output/maxent/paper_plots/prob_by_depth.png", width = 210, height = 297, units = "mm", dpi = 600)
```

# sample probabily maps

have moved sample files into a separate folder
march,  june, sep, dec
11.7737, 61.1128, 199.79. 628.026

```{r}
library(sf)
library(purrr)
library(raster)
library(lubridate)
library(tidyverse)

proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

s <- st_read("../data/env/landp/land-wgs84v2.shp") %>% 
  st_set_crs(proj)

mcp <- st_read("../output/bio/minimum_convex_poly/mcp-wgs84.shp")

filelist <- list.files("../data/plot2-probmaps/",pattern="avg_cloglog*",recursive = TRUE,include.dirs = TRUE)

x <- stack(file.path("../data/plot2-probmaps/",filelist)) %>% 
  mask(mcp)

names(x) <- filelist

df <- x %>% 
  rasterToPoints() %>% 
  as.data.frame() %>% 
  gather(key = "variable", value = "value",-x,-y) %>% 
  mutate(variable=gsub(".asc","",variable),
         month=factor(months(parse_date_time(gsub("\\..*$", "",variable),"%m")),
                      ordered = TRUE,
                      levels=months(parse_date_time(1:12,"%m"))),
         depth=as.numeric(gsub("^.*_", "",variable)))

#remi code stops here

df$month <- str_sub(df$variable, start = 16, end = 18)

CapStr <- function(y) {
  c <- strsplit(y, " ")[[1]]
  paste(toupper(substring(c, 1,1)), substring(c, 2),
      sep="", collapse=" ")
}
df$month <- sapply(df$month, CapStr)

#remi code starts here

depthlab <- paste(unique(df$depth),"m")[c(1,3,2,4)]
names(depthlab) <- unique(df$depth)[c(1,3,2,4)]

ggplot(df)+
  geom_sf(data=mcp,fill='grey')+
  geom_raster(aes(x,y,fill=value))+
  geom_sf(data=mcp,fill="transparent")+
  scale_fill_viridis(option="magma",
                       direction=-1, limits = c(0,1), breaks = c(0, 0.25, 0.5, 0.75, 1),
                       guide=guide_colourbar(barwidth=0.5, barheight = 15, title = NULL, draw.ulim = FALSE, draw.llim = FALSE))+
  facet_grid(depth~month,
             labeller = labeller(depth=depthlab))+
  theme_classic()+
  scale_x_continuous(breaks = seq(-70,-50,10))+
  theme(legend.position = "right",
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
        strip.text.x = element_text(face = "bold", size = 12),
        strip.text.y = element_text(face = "bold", size = 12))

ggsave("../output/maxent/paper_plots/test-overview.png", width = 210, height = 297, units = "mm", dpi = 600)
```

#2d maps - max kappa thresholds

this creates binary maps for each layer based on the maxkappa threshold
```{r}
binprob <- function(x){
  ifelse(x > 0 & x >= mk, 1,
  ifelse(x <= 0, 0, NA))
}

#get the maxkap value
alst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)

for (f in 1:length(alst)){
  mth <- str_sub(alst[f], start = 35, end = 37)
  af <- read.csv(alst[f], header = TRUE)
  af <- tail(af, n = 1)
  mk <- af$test_maxkap

#create binary maps per depth layer
  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/hotspot_mapping/prob_by_depth/", mth, "_maxkap_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/hotspot_mapping/prob_by_depth/", mth, "_maxkap_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }
}

```

then create an summary map....
```{r >maxkap prob monthly 2D representation}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mthna <- mthord[m]
  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_maxkap_.*.asc$"), full.names = TRUE)
  binstack <- stack(binaryprob_list)
  stacksum <- sum(binstack, na.rm = TRUE)
  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_maxkap_2d.asc"), overwrite=TRUE)
  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_maxkap_2d.png")))
  dev.off()
}

```

now create summary maps where values are binary - not counts of cells
```{r}
binprob <- function(x){
  ifelse(x > 0, 1,
  ifelse(x <= 0, 0, NA))
}

  clogasc_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = "*_maxkap_2d.asc", full.names = TRUE)
  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- str_sub(clogasc_list[cum], start = 46, end = -8)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }
```

now create a table to see how many cells there are >maxkap in each month (for editing plots)

```{r count cells >maxkap}

binaryprob_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = ".*_maxkap_2d.asc", full.names = TRUE)
sum2d <- data.frame(matrix(ncol = 2, nrow = 10))
sum2d <- setNames(sum2d, c("month", "count"))

for (r in 1:length(binaryprob_list)){
  ras <- raster(binaryprob_list[r])
  rnm <- names(ras)
  sum2d$month[r] <- gsub("_maxkap_2d", "", x = rnm)
  sum2d$count[r] <- cellStats(ras, sum, asSample = TRUE)
}

#order table from march to december
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
sum2d <- sum2d[match(mthord, sum2d$month), ]

#get proportion
tcs <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE) #file showing how many cells per depth
#tcs <- tcs[-c(1:3), ] #because the top three rows are deeper than the capelin obs...
totc <- sum(tcs$no_cells)

for (r in 1:nrow(sum2d)){
  sum2d$proportion[r] <- (sum2d$count[r]/totc)*100 #this actually shows percentage
}

write.csv(sum2d, "../output/maxent/paper_plots/cell_count_maxkap.csv", row.names = FALSE)

```

```{r plot maxkap 2d}
library(sf)
library(purrr)
library(raster)
library(lubridate)
library(tidyverse)

proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")

s <- st_read("../data/plot3-2d/landp/land-wgs84v2.shp") %>% 
  st_set_crs(proj) %>% 
  st_intersection(mcp)

filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", include.dirs = TRUE)

x <- stack(file.path("../output/maxent/paper_plots/hotspot_mapping/",filelist)) %>% 
  mask(mcp) %>% 
  rasterToPoints() %>% 
  as.data.frame() %>% 
  gather(key = "variable", value = "value",-x,-y) %>% 
  mutate(month=factor(months(parse_date_time(gsub("_.*$", "",variable),"%m")),
                      ordered = TRUE,
                      levels=months(parse_date_time(1:12,"%m"))),
         value=ifelse(value==1,">0.5","<0.5"))

df <- x %>% 
  group_by(month) %>% 
  summarize(total=round(sum(value==">0.5")/n(),2))

#quickest way might be just to inject my csv file with the correct values here....
df2 <- read.csv("../output/maxent/paper_plots/cell_count_maxkap.csv", header = TRUE)
df$total <- df2$proportion
df$total <- format(round(df$total, digits = 2))
#colnames(df) <- c("month", "cellc", "total")

ggplot(x)+
  geom_raster(aes(x,y,fill=value))+
  scale_fill_manual(values=c('light yellow','black'))+
  geom_sf(data=s,fill='grey',colour='transparent')+
  geom_sf(data=mcp,fill="transparent")+
  geom_text(data=df,x=-50,y=67,label=df$total)+
  facet_wrap(~month,nrow=2)+#,
  theme_classic()+
  scale_x_continuous(breaks = seq(-70,-50,10))+
  theme(legend.position = "none",
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(angle=45,vjust=1,hjust=1),
        strip.background = element_blank(),
        strip.text.x = element_text(face = "bold"))

ggsave("../output/maxent/paper_plots/maxkap_2d.png", height = 210, width = 297, units = "mm", dpi = 600)

```

get data to plot maxkap binary by depth... 
```{r maxkap by depth}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

binprob <- function(x){
  ifelse(x >0,  1,
  ifelse(x == -9999, 0, NA))
}

for (m in 1:length(mthord)){
  mth <- mthord[m]
  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_maxkap.*asc"), full.names = TRUE)
  binaryavgstack <- stack(ascdepth)
  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
  stackstathigh[ ,1] <- c(1:47)
  colnames(stackstathigh) <- c("asclayer", "prob_bin")

  for (lyr in 1:nlayers(binaryavgstack)){
    ascname <-  str_sub(ascdepth[lyr], 67, -5)
    dlay <- binaryavgstack[[lyr]]
    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
    lyrsum <- cellStats(dlay, sum)
    stackstathigh$prob_bin[lyr] <- lyrsum
    stackstathigh$asclayer[lyr] <- ascname

  }
  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_cellcount_bydepth_avg.csv"), row.names = FALSE)
}

```

think about how to get proportions...

```{r percentage cells above and below maxkap}
mthord <- c("mar",  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
cumst1 <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE)

for (m in 1:length(mthord)){
  mth <- mthord[m]
  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
  for (r in 1:nrow(a)){
    # a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
    # d1 <- str_sub(a$Depth[r], end = 4)
    d1 <- a$asclayer[r]
	  for (ro in 1:nrow(cumst1)){
      #d2 <- str_sub(cumst1$depth[ro], end = 4)
	    d2 <- cumst1$depth[ro]
      if (d1 == d2){
          a$total_cells[r] <- cumst1$no_cells[ro]
      }
	  }
    a$p_cells[r] <- (a$prob_bin[r]/a$total_cells[r])*100
    a$Probability[r] <- 1
  }
    #a$asclayer <- as.numeric(a$asclayer)

  a2 <- as.data.frame(a$asclayer)
  a2$total_cells <- a$total_cells
  a2$prob_bin <- a$total_cells - a$prob_bin
  a2$p_cells <- (a2$prob_bin/a2$total_cells)*100
  a2$Probability <- 0
  colnames(a2)[1] <- "asclayer"

  b <- rbind(a, a2)
  colnames(b)[1] <- "Depth"

  write.csv(b, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_dis_prob_cells.csv"), row.names = FALSE)
}

```

based on Remi's code - maxkappa derrived habitat by depth
```{r maxkappa habitat by depth}
require(data.table)
require(purrr)
require(lubridate)
require(tidyverse)
require(viridis)

filenames <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "_maxkap_dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)

all <- map_dfr(filenames, function(x) fread(x) %>% 
                       mutate(Month=substr(basename(x),1,3))) %>% 
  mutate(Depth = factor(round(Depth,1)),
         Depth = factor(Depth,ordered = TRUE, levels = rev(levels(Depth))),
         Month=factor(months(parse_date_time(Month,"%m")),
                      ordered = TRUE,
                      levels=months(parse_date_time(1:12,"%m"))))

#create a list of depths so you can set breaks for the x-axis (which is showing as the y-axis...)
dlst <- unique(all$Depth)
dlst <- as.character(dlst)
dlst <- as.numeric(dlst)
dlst <- sort(dlst)

#make probability a factor - sam added this
all$Probability <- as.factor(all$Probability)

#make the plot
  ggplot(all) +
  geom_col(aes(Depth, p_cells, fill = Probability),width=1) +
  coord_flip()+
  facet_wrap(~Month,ncol=2)+
  labs(y = "% Ocean at Depth", x = "Depth (meters)")+
  scale_fill_manual(aesthetics = "fill", values=c('light yellow','black'), name = "Presence")+
  theme_classic()+
  theme(legend.position = "bottom", axis.text.y = element_text(size=7), strip.background = element_blank()) +
  scale_x_discrete(breaks = c(dlst[1], dlst[5], dlst[10], dlst[15], dlst[20], dlst[25], dlst[30], dlst[35], dlst[40], dlst[47]))

ggsave("../output/maxent/paper_plots/maxkap_prob_by_depth.png", width = 210, height = 297, units = "mm", dpi = 600)
```

#get max and min depth layers based on max kappa
```{r maxkap min max depth}
kapd <- list.files("../output/maxent/paper_plots/prob_by_depth/" , pattern = "_maxkap_dis_prob_cells.csv", full.names = TRUE)
kapmmd <- data.frame(mth = character(10), min = numeric(10), max = numeric(10), range = numeric(10))
kapmmd$mth <- as.character(kapmmd$mth)

for(k in 1:length(kapd)){
  kapmmd$mth[k] <- str_sub(kapd[k], start = 44, end = 46)
  kd <- read.csv(kapd[k])
  kd <- subset(kd, Probability == 1) #just keep the presence cells
  kd <- subset(kd, prob_bin > 0) # get the depths where there are presence cells
  kd <- kd[order(kd[, 1]),] #order by depth 
  kapmmd$min[k] <- kd$Depth[1] #minimum depth of presence cell
  kapmmd$max[k] <- tail(kd$Depth, n = 1) #maximum depth of presence cell
}

kapmmd$range <- kapmmd$max - kapmmd$min #get depth range
write.csv(kapmmd, "../output/maxent/paper_plots/hotspot_mapping/kappa-min-max-depth.csv")
```

#get lat long position of most northely, southerly, western, and eastern cell centroid (max kap)

```{r maxkap cell centroid}
filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", full.names = TRUE)
lonlat <- data.frame(mth = character(10), west = numeric(10), east = numeric(10), north = numeric(10), south = numeric(10))
lonlat$mth <- as.character(lonlat$mth)

for (k in 1:length(filelist)){
  ras <- raster(filelist[k])
  crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" #the projection of the raster
  lonlat$mth[k] <- str_sub(filelist[k], start = 46, end = 48)
  cellloc <- rasterToPoints(ras, fun=function(x){x==1}) #make raster to point only for cells with the value of 1 (i.e. presence cells)
  cellloc <- as.data.frame(cellloc) # x = lon, y = lat
  cellloc <- cellloc[order(cellloc$x), ] #order dataframe by lon from low (west) to high (east)
  lonlat$west[k] <- cellloc$x[1] #most western cell centroid
  lonlat$east[k] <- tail(cellloc$x, n = 1) #most eastern cell centroid
  cellloc <- cellloc[order(cellloc$y), ] #order dataframe by lat from low (south) to high (north)
  lonlat$south[k] <- cellloc$y[1] #most southern cell centroid
  lonlat$north[k] <- tail(cellloc$y, n = 1) #most northern cell centroid
}

write.csv(lonlat, "../output/maxent/paper_plots/hotspot_mapping/kappa-nswe.csv")

```

#get lat long position of most northely, southerly, western, and eastern cell centroid for oct, june, aug, may (to remove northern spots)

```{r}
filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", full.names = TRUE)
lonlat <- data.frame(mth = character(10), west = numeric(10), east = numeric(10), north = numeric(10), south = numeric(10))
lonlat$mth <- as.character(lonlat$mth)

for (k in 1:length(filelist)){
  ras <- raster(filelist[k])
  crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" #the projection of the raster
  lonlat$mth[k] <- str_sub(filelist[k], start = 46, end = 48)
  cellloc <- rasterToPoints(ras, fun=function(x){x==1}) #make raster to point only for cells with the value of 1 (i.e. presence cells)
  cellloc <- as.data.frame(cellloc) # x = lon, y = lat
  cellloc <- subset(cellloc, y <55) #remove all the northern points
  cellloc <- cellloc[order(cellloc$y), ] #order dataframe by lat from low (south) to high (north)
  lonlat$south[k] <- cellloc$y[1] #most southern cell centroid
  lonlat$north[k] <- tail(cellloc$y, n = 1) #most northern cell centroid
  cellloc <- cellloc[order(cellloc$x), ] #order dataframe by lon from low (west) to high (east)
  lonlat$west[k] <- cellloc$x[1] #most western cell centroid
  lonlat$east[k] <- tail(cellloc$x, n = 1) #most eastern cell centroid

}

write.csv(lonlat, "../output/maxent/paper_plots/hotspot_mapping/kappa-nswe_nonorth.csv")

```

#background point selection plots

```{r}
test_back_df <- read.csv("../data/env/background_point_check/201406_allbackgroundpoints.csv", header = TRUE)
```

ok now create the first lot of random for 2014_06

```{r}
testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
```

make long data
```{r}
bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
bkpt$variable <- as.character(bkpt$variable)
bkpt$no_points <- as.character(bkpt$no_points)

bkpt$no_points[1:10000] <- "10,000"
bkpt$no_points[10001:30000] <- "20,000"
bkpt$no_points[30001:80000] <- "50,000"
bkpt$no_points[80001:180000] <- "100,000"
bkpt$no_points[180001:370000] <- "190,000"
bkpt$variable <- "temp"
bkpt$value[1:10000] <- testbk10000$temp_depth
bkpt$value[10001:30000] <- testbk20000$temp_depth
bkpt$value[30001:80000] <- testbk50000$temp_depth
bkpt$value[80001:180000] <- testbk100000$temp_depth
bkpt$value[180001:370000] <- testbk190000$temp_depth

bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt2$variable <- as.character(bkpt2$variable)
bkpt2$no_points[1:10000] <- "10,000"
bkpt2$no_points[10001:30000] <- "20,000"
bkpt2$no_points[30001:80000] <- "50,000"
bkpt2$no_points[80001:180000] <- "100,000"
bkpt2$no_points[180001:370000] <- "190,000"
bkpt2$variable <- "salinity"
bkpt2$value[1:10000] <- testbk10000$salinity_depth
bkpt2$value[10001:30000] <- testbk20000$salinity_depth
bkpt2$value[30001:80000] <- testbk50000$salinity_depth
bkpt2$value[80001:180000] <- testbk100000$salinity_depth
bkpt2$value[180001:370000] <- testbk190000$salinity_depth

bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt3$variable <- as.character(bkpt3$variable)
bkpt3$no_points[1:10000] <- "10,000"
bkpt3$no_points[10001:30000] <- "20,000"
bkpt3$no_points[30001:80000] <- "50,000"
bkpt3$no_points[80001:180000] <- "100,000"
bkpt3$no_points[180001:370000] <- "190,000"
bkpt3$variable <- "o2"
bkpt3$value[1:10000] <- testbk10000$o2_depth
bkpt3$value[10001:30000] <- testbk20000$o2_depth
bkpt3$value[30001:80000] <- testbk50000$o2_depth
bkpt3$value[80001:180000] <- testbk100000$o2_depth
bkpt3$value[180001:370000] <- testbk190000$o2_depth

bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt4$variable <- as.character(bkpt4$variable)
bkpt4$no_points[1:10000] <- "10,000"
bkpt4$no_points[10001:30000] <- "20,000"
bkpt4$no_points[30001:80000] <- "50,000"
bkpt4$no_points[80001:180000] <- "100,000"
bkpt4$no_points[180001:370000] <- "190,000"
bkpt4$variable <- "chl"
bkpt4$value[1:10000] <- testbk10000$chl_surface
bkpt4$value[10001:30000] <- testbk20000$chl_surface
bkpt4$value[30001:80000] <- testbk50000$chl_surface
bkpt4$value[80001:180000] <- testbk100000$chl_surface
bkpt4$value[180001:370000] <- testbk190000$chl_surface

bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
bk$no_points <- as.factor(bk$no_points)
```

```{r}
ggplot(bk, aes(x = value)) +
    stat_density(aes(group = no_points, color = no_points),position="identity",geom="line") +
  facet_wrap(~variable,ncol=2)

```

problem here is that i don't want the axis shared

fiddle with the dataset again
```{r}
t <- subset(bk, variable == "temp")
t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
s <- subset(bk, variable == "salinity")
s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
o <- subset(bk, variable == "o2")
o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
c <- subset(bk, variable == "chl")
c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))

```

plot
```{r}
p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))

p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")

ggsave("../output/maxent/paper_plots/bkpt_201406.png", width = 210, height = 297, units = "mm", dpi = 600)

```

200710

```{r}
test_back_df <- read.csv("../data/env/background_point_check/200710_allbackgroundpoints.csv", header = TRUE)
```

ok now create the first lot of random for 200710

```{r}
testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
```

make long data
```{r}
bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
bkpt$variable <- as.character(bkpt$variable)
bkpt$no_points <- as.character(bkpt$no_points)

bkpt$no_points[1:10000] <- "10,000"
bkpt$no_points[10001:30000] <- "20,000"
bkpt$no_points[30001:80000] <- "50,000"
bkpt$no_points[80001:180000] <- "100,000"
bkpt$no_points[180001:370000] <- "190,000"
bkpt$variable <- "temp"
bkpt$value[1:10000] <- testbk10000$temp_depth
bkpt$value[10001:30000] <- testbk20000$temp_depth
bkpt$value[30001:80000] <- testbk50000$temp_depth
bkpt$value[80001:180000] <- testbk100000$temp_depth
bkpt$value[180001:370000] <- testbk190000$temp_depth

bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt2$variable <- as.character(bkpt2$variable)
bkpt2$no_points[1:10000] <- "10,000"
bkpt2$no_points[10001:30000] <- "20,000"
bkpt2$no_points[30001:80000] <- "50,000"
bkpt2$no_points[80001:180000] <- "100,000"
bkpt2$no_points[180001:370000] <- "190,000"
bkpt2$variable <- "salinity"
bkpt2$value[1:10000] <- testbk10000$salinity_depth
bkpt2$value[10001:30000] <- testbk20000$salinity_depth
bkpt2$value[30001:80000] <- testbk50000$salinity_depth
bkpt2$value[80001:180000] <- testbk100000$salinity_depth
bkpt2$value[180001:370000] <- testbk190000$salinity_depth

bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt3$variable <- as.character(bkpt3$variable)
bkpt3$no_points[1:10000] <- "10,000"
bkpt3$no_points[10001:30000] <- "20,000"
bkpt3$no_points[30001:80000] <- "50,000"
bkpt3$no_points[80001:180000] <- "100,000"
bkpt3$no_points[180001:370000] <- "190,000"
bkpt3$variable <- "o2"
bkpt3$value[1:10000] <- testbk10000$o2_depth
bkpt3$value[10001:30000] <- testbk20000$o2_depth
bkpt3$value[30001:80000] <- testbk50000$o2_depth
bkpt3$value[80001:180000] <- testbk100000$o2_depth
bkpt3$value[180001:370000] <- testbk190000$o2_depth

bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt4$variable <- as.character(bkpt4$variable)
bkpt4$no_points[1:10000] <- "10,000"
bkpt4$no_points[10001:30000] <- "20,000"
bkpt4$no_points[30001:80000] <- "50,000"
bkpt4$no_points[80001:180000] <- "100,000"
bkpt4$no_points[180001:370000] <- "190,000"
bkpt4$variable <- "chl"
bkpt4$value[1:10000] <- testbk10000$chl_surface
bkpt4$value[10001:30000] <- testbk20000$chl_surface
bkpt4$value[30001:80000] <- testbk50000$chl_surface
bkpt4$value[80001:180000] <- testbk100000$chl_surface
bkpt4$value[180001:370000] <- testbk190000$chl_surface

bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
bk$no_points <- as.factor(bk$no_points)
```

fiddle with the dataset again
```{r}
t <- subset(bk, variable == "temp")
t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
s <- subset(bk, variable == "salinity")
s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
o <- subset(bk, variable == "o2")
o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
c <- subset(bk, variable == "chl")
c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))

```

plot
```{r}
p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))

p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")

ggsave("../output/maxent/paper_plots/bkpt_200710.png", width = 210, height = 297, units = "mm", dpi = 600)

```
199902

```{r}
test_back_df <- read.csv("../data/env/background_point_check/199902_allbackgroundpoints.csv", header = TRUE)
```

ok now create the first lot of random for 200710

```{r}
testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
```

make long data
```{r}
bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
bkpt$variable <- as.character(bkpt$variable)
bkpt$no_points <- as.character(bkpt$no_points)

bkpt$no_points[1:10000] <- "10,000"
bkpt$no_points[10001:30000] <- "20,000"
bkpt$no_points[30001:80000] <- "50,000"
bkpt$no_points[80001:180000] <- "100,000"
bkpt$no_points[180001:370000] <- "190,000"
bkpt$variable <- "temp"
bkpt$value[1:10000] <- testbk10000$temp_depth
bkpt$value[10001:30000] <- testbk20000$temp_depth
bkpt$value[30001:80000] <- testbk50000$temp_depth
bkpt$value[80001:180000] <- testbk100000$temp_depth
bkpt$value[180001:370000] <- testbk190000$temp_depth

bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt2$variable <- as.character(bkpt2$variable)
bkpt2$no_points[1:10000] <- "10,000"
bkpt2$no_points[10001:30000] <- "20,000"
bkpt2$no_points[30001:80000] <- "50,000"
bkpt2$no_points[80001:180000] <- "100,000"
bkpt2$no_points[180001:370000] <- "190,000"
bkpt2$variable <- "salinity"
bkpt2$value[1:10000] <- testbk10000$salinity_depth
bkpt2$value[10001:30000] <- testbk20000$salinity_depth
bkpt2$value[30001:80000] <- testbk50000$salinity_depth
bkpt2$value[80001:180000] <- testbk100000$salinity_depth
bkpt2$value[180001:370000] <- testbk190000$salinity_depth

bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt3$variable <- as.character(bkpt3$variable)
bkpt3$no_points[1:10000] <- "10,000"
bkpt3$no_points[10001:30000] <- "20,000"
bkpt3$no_points[30001:80000] <- "50,000"
bkpt3$no_points[80001:180000] <- "100,000"
bkpt3$no_points[180001:370000] <- "190,000"
bkpt3$variable <- "o2"
bkpt3$value[1:10000] <- testbk10000$o2_depth
bkpt3$value[10001:30000] <- testbk20000$o2_depth
bkpt3$value[30001:80000] <- testbk50000$o2_depth
bkpt3$value[80001:180000] <- testbk100000$o2_depth
bkpt3$value[180001:370000] <- testbk190000$o2_depth

bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt4$variable <- as.character(bkpt4$variable)
bkpt4$no_points[1:10000] <- "10,000"
bkpt4$no_points[10001:30000] <- "20,000"
bkpt4$no_points[30001:80000] <- "50,000"
bkpt4$no_points[80001:180000] <- "100,000"
bkpt4$no_points[180001:370000] <- "190,000"
bkpt4$variable <- "chl"
bkpt4$value[1:10000] <- testbk10000$chl_surface
bkpt4$value[10001:30000] <- testbk20000$chl_surface
bkpt4$value[30001:80000] <- testbk50000$chl_surface
bkpt4$value[80001:180000] <- testbk100000$chl_surface
bkpt4$value[180001:370000] <- testbk190000$chl_surface

bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
bk$no_points <- as.factor(bk$no_points)
```

fiddle with the dataset again
```{r}
t <- subset(bk, variable == "temp")
t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
s <- subset(bk, variable == "salinity")
s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
o <- subset(bk, variable == "o2")
o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
c <- subset(bk, variable == "chl")
c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))

```

plot
```{r}
p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))

p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")

ggsave("../output/maxent/paper_plots/bkpt_199902.png", width = 210, height = 297, units = "mm", dpi = 600)

```

# variable maps
- temp
- chl 
- o2
- sal
- 4 different months
- relationship of variables to each other
- pick a depth (say the 11.7737 m)

months are mar, jun, sep, dec
pick 3 diff years say 1998, 2006, 2013